### 2018-02-12 from_chead/ballerDepHeterogen/ballerDepHeterogenScripts/Prepforhydra_rmarkdown_20180122.Rmd
updates including hydra prep scripts

---
title: "Prep_for_hyrda_with_matching"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
The purpose of this script is to read in CNB data, subset it, merge it, and write files to use in Hydra.  Before doing hydra, however, we must address the vastly different ages, maternal education and race between depressed and control groups.  Unfortunately, the matching removes a high number of our patients.

```{r subset, echo = TRUE}
library(visreg)
library(mgcv)
library(tableone)
library(dplyr)
library(plm)
library(MatchIt)

#read in csvs
demographics <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_demographics_go1_20161212.csv
cnb_scores <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_cnb_zscores_fr_20170202.csv
health <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_health_20170405.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_health_20170405.csv
psych_summary <- read.csv("/Users/eballer/BBL/from_chead/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv", header = TRUE, sep = ",") #from /data/joy/BBL/projects/ballerDepHeterogen/data/n9498_goassess_psych_summary_vars_20131014.csv

#remove people with NA for race, age, or sex.  START WITH N = 9498
demographics_noNA_race <- demographics[!is.na(demographics$race),] #everyone has a race, N = 9498
demographics_noNA_race_age <- demographics_noNA_race[!is.na(demographics_noNA_race$ageAtClinicalAssess1),] # 86 people do not have age at clinical assessment.  N = 9412
demographics_noNA_race_age_sex <- demographics_noNA_race_age[!is.na(demographics_noNA_race_age$sex),] #everyone has a sex, N = 9412
demographics_noNA_race_age_andCNBage_sex <- demographics_noNA_race_age_sex[!is.na(demographics_noNA_race_age_sex$ageAtCnb1),] #6 people do not have ageAtCnb1, N = 9406

#remove people with NA for depression or total psych score, START WITH N = 9498
psych_summary_no_NA_dep <- psych_summary[!is.na(psych_summary$smry_dep),] #take out those with NA for depression, 87 people N = 9411
psych_summary_no_NA_dep_and_smry_psych_overall <- psych_summary_no_NA_dep[!is.na(psych_summary_no_NA_dep$smry_psych_overall_rtg),] #take out those with NA for overall psych rtg, no additional people lost, N = 9411

#merge the csvs
#merge demographics and cnb #this is if we want to include people without full demographic data
dem_cnb <- merge(demographics_noNA_race_age_andCNBage_sex, cnb_scores, by = "bblid") #merge demographics and cnb, N = 9406
psych_health <- merge(psych_summary_no_NA_dep_and_smry_psych_overall, health, by = "bblid") #merge psych and health, N = 9411
dem_cnb_psych_health_merged <- merge (dem_cnb, psych_health, by = "bblid") #merge all 4 csvs, lost 1 person [134716] (had demographics, but no psych ratings): N = 9405

#make subsets
subset_just_dep_and_no_medicalratingExclude <- subset.data.frame(dem_cnb_psych_health_merged, (medicalratingExclude == 0) & (smry_dep == 4), select = "bblid") #subset people who were not medically excluded and who are depressed, N = 776
subset_no_psych_no_medicalratingExclude <- subset.data.frame(dem_cnb_psych_health_merged, (medicalratingExclude == 0) & (smry_psych_overall_rtg < 4), select = "bblid") #subset people who are psychiatrically healthy, N = 2508
subset_dep_or_no_psych_and_no_medicalratingExclude <- subset.data.frame(dem_cnb_psych_health_merged, (medicalratingExclude == 0) & ((smry_dep == 4) | (smry_psych_overall_rtg <4))) #subset including both depressed and healthies, good for regressions, N = 3284

#would binarize depression smry score to -1 (less than 4, not depressed) and 1 (score 4 , depressed)
dep_binarized <- ifelse(subset_dep_or_no_psych_and_no_medicalratingExclude$smry_dep == 4, 1, -1)
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED <- cbind(subset_dep_or_no_psych_and_no_medicalratingExclude, dep_binarized) #N = 3284

#make depression and gender into factor scores
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$dep_binarized <- as.factor(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$dep_binarized)
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex <- as.factor(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex)

#divide ageAtCNB by 12 for age
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$age_in_years <- subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$ageAtCnb1/12

#age demeaned and squared, from Toni
subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$ageSq <- as.numeric(I(scale(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$age_in_years, scale = FALSE, center = TRUE)^2))

#Subset only variables needed for hydra analysis 
#(BBLID, cognitive variables, depression), also do by males(1555)/females(1729) separately
subset_bblidAndCog_features <- data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[14:39], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[77]))
subset_bblidAndCog_features_males <- subset.data.frame(data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[14:39], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[77])), subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex == 1)
subset_bblidAndCog_features_females <- subset.data.frame(data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[14:39], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[77])), subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex == 2)

#subset of covariates (BBLID, sex, age in years), also do by males/females
subset_bblidAndCovariates <- data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1:2], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[78]))
subset_bblidAndCovariates_males <- subset.data.frame(data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1:2], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[78])), subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex == 1)
subset_bblidAndCovariates_females <- subset.data.frame(data.frame(cbind(subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[1:2], subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED[78])), subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED$sex == 2)

#save files for hyd
write.csv(subset_bblidAndCog_features, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_all_gender/Features.csv", row.names = FALSE, quote = FALSE)
write.csv(subset_bblidAndCovariates, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_all_gender/Covariates.csv", row.names = FALSE, quote = FALSE)
write.csv(subset_bblidAndCog_features_males, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_males/Features.csv", row.names = FALSE, quote = FALSE)
write.csv(subset_bblidAndCovariates_males, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_males/Covariates.csv", row.names = FALSE, quote = FALSE)
write.csv(subset_bblidAndCog_features_females, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_females/Features.csv", row.names = FALSE, quote = FALSE)
write.csv(subset_bblidAndCovariates_females, file="/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydra_females/Covariates.csv", row.names = FALSE, quote = FALSE)

```

## Including Plots

###

```{r pressure, echo=TRUE}
#match with matchit (starting n = 3284, males = 1555, females 1729)
data.unmatched = subset_dep_or_no_psych_and_no_medicalratingExclude_DEPBINARIZED
data.unmatched$unmatchedRows =rownames(data.unmatched)
dataset = data.unmatched
# Some preprocessing
dataset = dplyr::select(dataset, sex, age_in_years, ageSq, medu1, race, dep_binarized, unmatchedRows)
#dataset = dplyr::filter(dataset, !is.na(group))
# Dep: 1, Health = 0 
dataset$dep_binarized = 1*(dataset$dep_binarized==1) 
#"male": 1, "female": 0
dataset$sex = 1*(dataset$sex==1)
# Remove subjects with NA for maternal edu, new N = 3256, males = 1539, females 1717
dataset <- dataset[!is.na(dataset$medu1),]

# Plot prematch
plot(dataset$age_in_years,jitter(dataset$medu1, factor=3), col=2*dataset$race+dataset$dep_binarized+1,pch=16, ylab="Maternal Edu", xlab="Age")
legend("bottomright",c("Non-white, non-depressed", "Non-white, depressed", "White, non-depressed", "White, depressed"),pch=rep(16, 4), col=c(1,2,3,4))

#GAM for propensity score
ps.model =gam(dep_binarized ~s(age_in_years) +s(medu1) + race + sex, data=dataset, family=binomial)
ps =exp(predict(ps.model))/(1 +exp(predict(ps.model)))
m.out <-matchit(dep_binarized ~ age_in_years, data=dataset, method="nearest", exact=c("race", "medu1", "sex"), distance="mahalanobis")
plot(m.out)

#return the matched dataset. N = 1518, males = 518, females = 1000
m.data <- match.data(m.out)

# Test for significant difference in age between groups
t.test(age_in_years~dep_binarized, data=m.data)
t.test(race~dep_binarized, data=m.data)
t.test(sex~dep_binarized, data=m.data)
t.test(medu1~dep_binarized, data=m.data)

# Re-plot
plot(m.data$age_in_years,jitter(m.data$medu1, factor=3), col=2*m.data$race+m.data$dep_binarized+1,pch=16, ylab="Maternal Edu", xlab="Age")
legend("bottomright",c("Non-white, non-depressed", "Non-white, depressed", "White, non-depressed", "White, depressed"),pch=rep(16, 4), col=c(1,2,3,4))

# Make the final matched data set
data.matched = data.unmatched[data.unmatched$unmatchedRows%in%m.data$unmatchedRows,]
data.matched$unmatchedRows = NULL

saveRDS(data.matched, file='/Users/eballer/BBL/from_chead/ballerDepHeterogen/results/hydraMatched_age_race_medu1_sex_20180115_matched.rds')
```