### 2020-10-29 lab_notebook.Rmd

---
title: "R Notebook for MSDepression Project"
output:
  html_notebook
---

#### 20201008

First, need to summarize data

1) Got data into csv format
  - Opened /Users/eballer/Documents/Penn T32/MS_Depression_Ted_Taki/Documents/*xls in excel and saved as csv
2)
```{move_data bash}
cd /Users/eballer/BBL/msdepression/data/
mkdir dac
cp /Users/eballer/Documents/Penn T32/MS_Depression_Ted_Taki/Documents/*csv .
```

3) In R, started /Users/eballer/BBL/msdepression/dac_summary.Rmd to summarize data

** during the last iteration of de-identifying the DAC, the PHQ-9 scores all vanished. Will need to ask for them to be re-added.
** also, number of people dropped by 50% (from 36277 to 16383)

#### 20201012

Continued with descriptives

  - Read in Melissa's data sheet.
  - subsetted EMPI and Accession # (since our EMPI's are deidentified)
  - Merged with my datasheet on Accession #
  - "PatientID == EMPI"

Merged Melissa's data with mine to get empis.

#### 20201013

[x] Figure out how many people with repeat mrns - to remove doubles. Will need to remove duplicates.

#### 20201027

Going back and forth a lot with DAC people, should have a new update soon. Big thing will be to parse PHQ2/9. We will get all of them but have to tie them to their appropriate scans