### 2020-11-05 dac_summary.Rmd
Add files via upload
- added details combining my and melissa's data. Made lots of changes to unique/duplicated code.

---
title: "Dac Summary"
output:
  html_document
---

## Dac summary
First run on: 20201008
Spreadsheet:

This script goes through data that has been returned from the DAC and summarizes it
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r summary}
#preprocessing - because these are sent

#Load original MS database from melissa
#load("/Users/eballer/BBL/msdepression/data/database_2019-03-04.RData")

#read in melissa's xls: n=30915
melissa_xls <- read.csv("/Users/eballer/BBL/msdepression/data/mri_file_upd_w_empi.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"))

#load our df
data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninpatientswithmsmaskdataaddedmoredxcodes.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL")) #n=42101, with 42044 unique accession codes

data2 <- read.csv("/Users/eballer/Documents/Penn T32/MS_Depression_Ted_Taki/Documents/investigatingdepressioninpatientswithmultiplesclerosis10-23medsinonerow (1).csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"))
#subset EMPI, accession codes, F-codes, Phq2 and phq9 #n = 29226, with 21378 unique accession codes

#drawn from Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninpatientswithmsmaskdataaddedmoredxcodes.csv, which was split into 2 b/c the spreadsheet was too big 
data3_p1 <- read.csv("/Users/eballer/BBL/msdepression/data/dac/20201029_msdep_pt1.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep = ",") # n = 28556 with 20871 unique accession codes

data3_p2 <- read.csv("/Users/eballer/BBL/msdepression/data/dac/20201029_msdep_pt2.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep = ",")

data3 <- rbind(data3_p1, data3_p2)
write.csv(x = data3, file = "/Users/eballer/BBL/msdepression/data/dac/msdep_20201029_merged.csv")

#set data_empi_acc_f_phq to data3
data_empi_acc_f_phq <- data3#subset(data, select = c("EMPI", "ACCESSION_NUM", "PHQ.2", "PHQ.9"))

#number of unique in melissa's
m_access_n <- length(unique(melissa_xls$ACCESSION_NUM)) #n = 29183

#number of unique in my group
eb_access_n <- length(unique(data_empi_acc_f_phq$ACCESSION_NUM)) #n = 21027

#number of overlap

melissa_peeps_in_eb_group <- sum(unique(melissa_xls$ACCESSION_NUM) %in% unique(data_empi_acc_f_phq$ACCESSION_NUM)) #n = 15137

eb_peeps_in_melissa_group <- sum(unique(data_empi_acc_f_phq$ACCESSION_NUM) %in% unique(melissa_xls$ACCESSION_NUM)) #n=15137

print(paste0("Number of Melissa's people in EB group : ", melissa_peeps_in_eb_group, " out of ", m_access_n, " unique accession numbers")) #suggesting that Melissa has a lot of people that don't have ms defined the same way we do

print(paste0("Number of EB's people in Melissa's group : ", eb_peeps_in_melissa_group, " out of ", eb_access_n, " unique accession numbers"))

overlap_group <- data_empi_acc_f_phq[which(unique(data_empi_acc_f_phq$ACCESSION_NUM) %in% unique(melissa_xls$ACCESSION_NUM)),] #n=15137

##########################
### some preprocessing ###
##########################

#change accession name for ease of merging with Rdata if using
#names(data_empi_acc_f_phq) <- gsub(pattern = "ACCESSION_NUM", replacement = "AccessionNumber", names(data_empi_acc_f_phq))

#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
data_empi_acc_f_phq[28:42] <- apply(X = data_empi_acc_f_phq[28:42], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)

#get empis of people who have wacky deidentified empis
db_empi_acc <- subset(melissa_xls, select = c("ACCESSION_NUM", "EMPI"))

#change EMPI to "patient ID"
names(db_empi_acc) <- gsub(pattern = "EMPI", replacement = "PatientID", names(db_empi_acc))
                      #subset(database, select = c("AccessionNumber", "PatientID"))

#merge with our database, PatientID is De-identified EMPI, n = 30799
data_empi_acc_f_phq_ptID <- merge(data_empi_acc_f_phq, db_empi_acc, by = "ACCESSION_NUM")

# number of unique scanners and scans
print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(unique(data_empi_acc_f_phq$PROC_CODE))))
print(paste0("Number of unique department names (ie. RAD HUP PCAM Ground): ",  length(unique(data_empi_acc_f_phq$DEPARTMENT_NAME))))
print(paste0("Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): ",  length(unique(data_empi_acc_f_phq$MODALITY))))

#phq-2;
print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2))))

#print histogram of phq-2s for people

hist(data_empi_acc_f_phq$PHQ.2[which(!is.na(data_empi_acc_f_phq$PHQ.2))], main = paste0("PHQ-2 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))), xlab = "PHQ-2", breaks = 8)

hist(data_empi_acc_f_phq_ptID$PHQ.2[which(!is.na(data_empi_acc_f_phq_ptID$PHQ.2))], main = paste0("PHQ-2 Histo in Melissa's subgroup #: n = ", length(which(!is.na(data_empi_acc_f_phq_ptID$PHQ.2)))), xlab = "PHQ-2", breaks = 8)

#phq-9
print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9))))

#print histogram of phq-9s for people
hist(data_empi_acc_f_phq$PHQ.9[which(!is.na(data_empi_acc_f_phq$PHQ.9))], main = paste0("PHQ-9 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.9)))), xlab = "PHQ-9", breaks = 30)

hist(data_empi_acc_f_phq_ptID$PHQ.9[which(!is.na(data_empi_acc_f_phq_ptID$PHQ.9))], main = paste0("PHQ-9 Histo In Melissa's group: n = ", length(which(!is.na(data_empi_acc_f_phq_ptID$PHQ.9)))), xlab = "PHQ-9", breaks = 30)

#both phq2 and 9

num_both_phq2_and_9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & !is.na(data_empi_acc_f_phq$PHQ.9)))
print(paste0("num people with both phq2 and 9 = ", num_both_phq2_and_9))

#looking at just people with phq-2 and 9
phq2_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.2)),] #n=5855
phq9_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.9)),] #n=606

#checking logic for order of subsetting - unique EMPI then ph2/9 == phq2/9 then EMPI subsetting - this works
unique_phq2_empis <- length(unique(phq2_subset$EMPI)) #n=1005
unique_phq9_empis <- length(unique(phq9_subset$EMPI)) #n=70

#age histogram
hist(data_empi_acc_f_phq$MRI_ENC_AGE[which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE))], main = paste0("Age Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")

#tsh
hist(data_empi_acc_f_phq$TSH[which(!is.na(data_empi_acc_f_phq$TSH))], main = paste0("TSH Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$TSH)))), xlab = "TSH")

#hemoglobin
hist(data_empi_acc_f_phq$hemoglobin[which(!is.na(data_empi_acc_f_phq$hemoglobin))], main = paste0("Hemoglobin Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$hemoglobin)))), xlab = "hemoglobin")

#folate
hist(data_empi_acc_f_phq$FOLATE[which(!is.na(data_empi_acc_f_phq$FOLATE))], main = paste0("Folate Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$FOLATE)))), xlab = "folate", breaks = 10)

#sex
hist(data_empi_acc_f_phq$sex_binarized[which(!is.na(data_empi_acc_f_phq$sex_binarized))], 
     main = paste0("Sex Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$sex_binarized))), "; M = ", length(which(data_empi_acc_f_phq$sex_binarized == 1)), "; F = ", length(which(data_empi_acc_f_phq$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

#race
hist(data_empi_acc_f_phq$race_binarized[which(!is.na(data_empi_acc_f_phq$race_binarized))], 
     main = paste0("Race Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$race))), "; White = ", length(which(data_empi_acc_f_phq$race_binarized == 1)), "; NW = ", length(which(data_empi_acc_f_phq$race_binarized == 2))),
     xlab = "white = 1, other = 2", 
     breaks = 2)

#f-codes - people usually have BOTH icd9/10 or neither
num_depression <- length(which(!is.na(data_empi_acc_f_phq$ICD9))) #n=11131
print(paste0("N with ICD 9/10 codes for depression = ", num_depression))

num_unique_EMPI_and_dep <- length(which(!is.na(data_empi_acc_f_phq$ICD9) & unique(data_empi_acc_f_phq$EMPI))) #n = 593
print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep))

num_phq2_and_dep <- length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.2))) #n = 4855
print(paste0("N with ICD 9/10 codes for depression and phq2 = ", num_phq2_and_dep))

num_phq9_and_dep <- length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.9))) #n=669
print(paste0("N with ICD 9/10 codes for depression and phq9 = ", num_phq9_and_dep))

#fcode histos
hist(data_empi_acc_f_phq$PHQ.2[which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.2))], 
     main = paste0("PHQ2 and Dep Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.2)))), xlab = paste0("PHQ2"))

hist(data_empi_acc_f_phq$PHQ.9[which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.9))], 
     main = paste0("PHQ9 and Dep Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$ICD9) & !is.na(data_empi_acc_f_phq$PHQ.9)))), xlab = paste0("PHQ9"))

############ 11/5/2020 #############
#########################
##### overlap group #####
#########################

#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
overlap_group$ACCESSION_NUM <- as.integer(overlap_group$ACCESSION_NUM)
overlap_group[28:42] <- apply(X = overlap_group[28:42], FUN = as.integer, MARGIN = 2)
overlap_group$sex_binarized <- ifelse(overlap_group$SEX == "MALE", 1, 2)
overlap_group$race_binarized <- ifelse(overlap_group$RACE == "WHITE", 1, 2)

#remove people who do not have phq2 or 9
overlap_group_with_phq2 <- overlap_group[which(!is.na(overlap_group$PHQ.2)),] #4864
overlap_group_with_phq9 <- overlap_group[which(!is.na(overlap_group$PHQ.9)),] #376

#subset people with unique scans (these are one time scanners)
overlap_group_no_duplicate_EMPI <- overlap_group[which(!duplicated(overlap_group$EMPI)),] #2675

overlap_group_with_phq2_no_duplicate_EMPI <- overlap_group_with_phq2[which(!duplicated(overlap_group_with_phq2$EMPI)),] #670

overlap_group_with_phq9_no_duplicate_EMPI <- overlap_group_with_phq9[which(!duplicated(overlap_group_with_phq9$EMPI)),] #44

#####take duplicated people and take first instance

#get duplicated people
overlap_group_with_duplicate_EMPI <- overlap_group[which(duplicated(overlap_group$EMPI)),]

overlap_group_with_phq2_w_duplicate_EMPI <- overlap_group_with_phq2[which(duplicated(overlap_group_with_phq2$EMPI)),]

overlap_group_with_phq9_w_duplicate_EMPI <- overlap_group_with_phq9[which(duplicated(overlap_group_with_phq9$EMPI)),]

#get match - i.e. first instance of each
overlap_group_with_duplicated_EMPI_first_instance <- overlap_group_with_duplicate_EMPI[match(unique(overlap_group_with_duplicate_EMPI$EMPI), overlap_group_with_duplicate_EMPI$EMPI),]

overlap_group_with_phq2_duplicated_EMPI_first_instance <- overlap_group_with_phq2_w_duplicate_EMPI[match(unique(overlap_group_with_phq2_w_duplicate_EMPI$EMPI), overlap_group_with_phq2_w_duplicate_EMPI$EMPI),]

overlap_group_with_phq9_duplicated_EMPI_first_instance <- overlap_group_with_phq9_w_duplicate_EMPI[match(unique(overlap_group_with_phq9_w_duplicate_EMPI$EMPI), overlap_group_with_phq9_w_duplicate_EMPI$EMPI),]

#combine groups
overlap_group_unique_EMPI_combined <- rbind(overlap_group_no_duplicate_EMPI, overlap_group_with_duplicated_EMPI_first_instance)

overlap_group_phq2_unique_empi_combined <- rbind(overlap_group_with_phq2_no_duplicate_EMPI, overlap_group_with_phq2_duplicated_EMPI_first_instance)

overlap_group_phq9_unique_empi_combined <- rbind(overlap_group_with_phq9_no_duplicate_EMPI, overlap_group_with_phq9_duplicated_EMPI_first_instance)

#summarize
print(paste0("N unique EMPIs in EB and Melissa overlap group = ", length(unique(overlap_group$EMPI)), " out of ", dim(overlap_group)[1])) #

print(paste0("N unique EMPIs in EB and Melissa overlap group w/phq2 = ", dim(overlap_group_phq2_unique_empi_combined)[1], " out of ", dim(overlap_group)[1])) #

print(paste0("N unique EMPIs in EB and Melissa overlap group w/phq9 = ", dim(overlap_group_phq9_unique_empi_combined)[1], " out of ", dim(overlap_group)[1])) #

#histos
hist(overlap_group_phq2_unique_empi_combined$PHQ.2, main = paste0("PHQ-2 Histo combined group w/unique EMPI vs all EMPI: n = ", dim(overlap_group_phq2_unique_empi_combined)[1], "/", length(which(!is.na(overlap_group$PHQ.2)))), xlab = "PHQ-2", breaks = 6)

#histos
hist(overlap_group_phq9_unique_empi_combined$PHQ.9, main = paste0("PHQ-9 Histo combined group w/unique EMPI vs all EMPI: n = ", dim(overlap_group_phq9_unique_empi_combined)[1], "/", length(which(!is.na(overlap_group$PHQ.9)))), xlab = "PHQ-9", breaks = 6)

#########################
### Erica's group  ####
#########################
data_e <- data3

#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
data_e$ACCESSION_NUM <- as.integer(data_e$ACCESSION_NUM)
data_e[28:42] <- apply(X = data_e[28:42], FUN = as.integer, MARGIN = 2)
data_e$sex_binarized <- ifelse(data_e$SEX == "MALE", 1, 2)
data_e$race_binarized <- ifelse(data_e$RACE == "WHITE", 1, 2)

#remove people who do not have phq2 or 9
data_e_with_phq2 <- data_e[which(!is.na(data_e$PHQ.2)),] #4864
data_e_with_phq9 <- data_e[which(!is.na(data_e$PHQ.9)),] #376

#subset people with unique scans (these are one time scanners)
data_e_no_duplicate_EMPI <- data_e[which(!duplicated(data_e$EMPI)),]

data_e_with_phq2_no_duplicate_EMPI <- data_e_with_phq2[which(!duplicated(data_e_with_phq2$EMPI)),]

data_e_with_phq9_no_duplicate_EMPI <- data_e_with_phq9[which(!duplicated(data_e_with_phq9$EMPI)),]

#####take duplicated people and take first instance

#get duplicated people
data_e_with_duplicate_EMPI <- data_e[which(duplicated(data_e$EMPI)),]

data_e_with_phq2_w_duplicate_EMPI <- data_e_with_phq2[which(duplicated(data_e_with_phq2$EMPI)),]

data_e_with_phq9_w_duplicate_EMPI <- data_e_with_phq9[which(duplicated(data_e_with_phq9$EMPI)),]

#get match - i.e. first instance of each
data_e_with_duplicated_EMPI_first_instance <- data_e_with_duplicate_EMPI[match(unique(data_e_with_duplicate_EMPI$EMPI), data_e_with_duplicate_EMPI$EMPI),]

data_e_with_phq2_duplicated_EMPI_first_instance <- data_e_with_phq2_w_duplicate_EMPI[match(unique(data_e_with_phq2_w_duplicate_EMPI$EMPI), data_e_with_phq2_w_duplicate_EMPI$EMPI),]

data_e_with_phq9_duplicated_EMPI_first_instance <- data_e_with_phq9_w_duplicate_EMPI[match(unique(data_e_with_phq9_w_duplicate_EMPI$EMPI), data_e_with_phq9_w_duplicate_EMPI$EMPI),]

#combine groups
data_e_unique_EMPI_combined <- rbind(data_e_no_duplicate_EMPI, data_e_with_duplicated_EMPI_first_instance)

data_e_phq2_unique_empi_combined <- rbind(data_e_with_phq2_no_duplicate_EMPI, data_e_with_phq2_duplicated_EMPI_first_instance)

data_e_phq9_unique_empi_combined <- rbind(data_e_with_phq9_no_duplicate_EMPI, data_e_with_phq9_duplicated_EMPI_first_instance)

#summarize
#print(paste0("N unique EMPIs in EB group = ", dim(data_e_unique_EMPI_combined)[1], " out of ", dim(data_e)[1])) #

print(paste0("N unique EMPIs in EB group = ", length(unique(data_e$EMPI)), " out of ", dim(data_e)[1]))

print(paste0("N unique EMPIs in EB group w/phq2 = ", dim(data_e_phq2_unique_empi_combined)[1], " out of ", dim(data_e)[1])) #

print(paste0("N unique EMPIs in EB group w/phq9 = ", dim(data_e_phq9_unique_empi_combined)[1], " out of ", dim(data_e)[1])) #

print(paste0("N in EB group w/phq2 = ", length(which(!is.na(data_e_no_duplicate_EMPI$PHQ.2))), " out of ", dim(data_e)[1])) #

print(paste0("N in EB group w/phq9 = ", length(which(!is.na(data_e_no_duplicate_EMPI$PHQ.9))), " out of ", dim(data_e)[1])) #

print(paste0("N unique EMPIs in Melissa's group = ", length(unique(melissa_xls$EMPI)), " out of ", dim(melissa_xls)[1]))

print(paste0("N in EB + Melissa overlap group w/depression ICD 9/10 = ", length(which(!is.na(overlap_group$ICD9))), " out of ", dim(overlap_group)[1])) #

print(paste0("N in EB & Melissa's group with dep ICD 9/10 and phq2 = ", length(which(!is.na(overlap_group_with_phq2$ICD9))), " out of ", dim(overlap_group)[1]))

print(paste0("N in EB & Melissa's group with dep ICD 9/10 and phq9 = ", length(which(!is.na(overlap_group_with_phq9$ICD9))), " out of ", dim(overlap_group)[1]))

#histos
hist(data_e_phq2_unique_empi_combined$PHQ.2, main = paste0("PHQ-2 Histo EB group w/unique EMPI vs all EMPI: n = ", dim(data_e_phq2_unique_empi_combined)[1], "/", length(which(!is.na(data_e$PHQ.2)))), xlab = "PHQ-2", breaks = 6)

#histos
hist(data_e_phq9_unique_empi_combined$PHQ.9, main = paste0("PHQ-9 Histo EB group w/unique EMPI vs all EMPI: n = ", dim(data_e_phq9_unique_empi_combined)[1], "/", length(which(!is.na(data_e$PHQ.9)))), xlab = "PHQ-9", breaks = 6)

```