### 2021-04-15 final_dac_summary_pull_20210119.Rmd
Add files via upload

---
title: "Dac summary 01/19/2021"
author: "Erica Baller"
date: "4/15/2021"
output: html_document
---

title: "Dac Summary"
output:
  html_document: default
  pdf_document: default
---

## Dac summary

Welcome to the MS project! This is a really exciting project dedicated to understanding psychopathology, particular depression in patients with MS. It is unqiue in a variety of ways, primarily that it uses clinical images that are research grade.

The team so far is big.
PIs- Ted Satterthwaite and Taki Shinohara
Senior Scientists - Azeez Adebimpe and Matt Cieslak
Research Specialist- Timothy Robert-Fitzgerald
Graduate students - Melissa Martin
Data Analysist - Sidney Covitz
Collaborators - Aaron Alexander-Bloch and Jenna Young
DAC Pull done by - Victoria Rautman(2020/2021) and Sunil Thomas (2018)

Date of FINAL Pull: 1/19/2021
Spreadsheet: /Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients.csv 
Summary of request to Victoria @ DAC:
---
9/2/2020

DAC Request
Non-funded IRB Approved Research  (enter IRB number): 843669
Criteria	Display?	Description / Exclusions / Limitations / Filters
MRN	Y	Unique identifier requested, but can be masked MRN
Visit ID	Y	
Patient Class(es)
Please select only which class(es) you will need.		x  Inpatient
 X   Outpatient
x  Emergency
Age or DOB ranges	Y	
Gender	Y	
Race	Y	
Department(s)
Provide department numbers not just names.	N	
Provider(s)
Provide ID’s not just names.	N	
Date(s)
Include in the specific range and date types (eg, admit, order, result)	1/1/2010-
12/31/20	All scans based on date of first treatment with glatiramer acetate
Procedure
Please include the specific procedure codes.  (ICD9/ICD10 is preferred for inpatient)	Y	88.91 (Magnetic Resonance Imaging of Brain)

Diagnosis
Please include the specific ICD-9/ICD-10 codes including all decimal points. Do not simply include ranges or wildcards.	Y	Multiple sclerosis (ICD-9: 340; ICD-10: G35), Depressive disorders (ICD-9: 296.99, 296.21, 296.22, 296.23, 296.24, 296.25, 296.26, 296.20, 296.31, 296.32, 296.33, 296.34, 296.35, 296.26, 296.30, 300.4, 293.83, 311, ICD-10: F32.0, F32.1, F32.2, F32.3, F32.4, F32.5, F32.6, F33.0, F33.1, F33.2, F33.41, F33.42, F33.9, F34.1, F06.31, F06.32, F06.34, F32.8, F32.9, major depressive disorder, persistent depressive disorder, depressive disorder due to another medical condition, other specified depressive disorder, unspecified depressive disorder)
Orders	N	
Medication
Please list as it is ordered within the UPHS EMR’s – medication id’s preferred	Y	Glatiramer Acetate (Copaxone)
Lab Result
Please list the lab as it is ordered within the UPHS EMR’s.	Y	
Other	Y	EDSS scores (if present)
Other	Y	All medications patients are receiving
Other	Y	All PACS accession numbers available
Other	Y	PHQ-2 and PHQ-9 if present
---
9/16/2020:
Per her note to me: We also specified that we wanted to pull images with ORIG_CPT values of 70551, 70552, 70553, MHDI, IMGMR0128, and we wanted scans with department ID 361, 547, 6004, or 6013. And I may have mentioned this before but the ICD9 and ICD10 codes for MS patients are 340 and G35, respectively

*For us, we are additionally interested in ALL ICD9/10 codes, ALL medications they are on at the time of scan, specific labs (as indicated), EDSS , PHQ-9 and 2 scores. The particular reason for this is that we are going to be looking at the relationship of brain imaging to depression in the MS group.

Mock columns: DE_PAT_ID	PAT_ID	EMPI	HUP_MRN	PAT_NM_WMRN	SEX	RACE	ETHNIC_GROUP	ORDER_ID	ORDERING_DTTM	BEGIN_EXAM_DTTM	END_EXAM_DTTM	TECH_USER_ID	TECHNOLOGIST	ACCESSION_NUM	PROC_ID	PROC_NAME	PROC_CODE	ORIG_CPT	PERFORMING_CSN_ID	COVERAGE_ID	FIN_CLASS_NAME	VISIT_EPM_ID	PERFORMING_DEP_ID	DEPARTMENT_NAME	MODALITY	ASSOCIATED_ICD9	ASSOCIATED_ICD10	PAT_AGE_AT_EXAM	MRI_ENC_AGE	CURRENT_AGE	All ICD-9 or 10 codes	All medication codes	wbc	rbc	hemoglobin	csf studies…	phq-9 score (and date)	phq-2 score (and date)	b12	folate 	TSH	RPR Vit D

---
 1/8/2021 email update (me to Victoria Rautman):
 What we realized is that the best way to define inclusion for the MS group is to focus on people who were seen at some point in neurology clinic.

If you’d be so kind, I’d be grateful for *hopefully the last pull. We would use the previous one, but instead of using the ICD 9/10 codes associated with the actual scan, we would use the location of clinic visits, and pull ALL scans, before or after diagnosis.

New updates:
 -Pull all the scans for patients who have ever been seen by “NEUROLOGY SOUTH PAVILLION” and “NEUROLOGY HUP” departments for clinic.

-Please include provider name associated with these department visits.

-Don’t worry about doing exclusions based on ICD9/10 codes for the scans themselves, please include all and we will sort through it ourselves

***Please also include vitamin D levels in addition to other labs***

The rest of the query would be the same as the most recent query you sent back.
---

Radiology sent images to Taki and group 4/2021

This script goes through data that has been returned from the DAC on 1/19 and summarizes it
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r summary}
#preprocessing - because these are sent

#load our df

#from Jan 2021 pull
data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")

ms_providers <- c("MARKOWITZ, CLYDE E.", "JACOBS, DINA A.", "WILLIAMSON, ERIC MICHAEL-LEE", "BERGER, JOSEPH ROBERT", "BERGER DO, JOSEPH", "PRUITT, AMY A.", "KOLSON, DENNIS L.", "CHAHIN, SALIM", "NARULA, SONA", "BAR-OR, AMIT")

data_empi_acc_f_phq <- data[which(data$Provider %in% ms_providers),] #n = 17067
table(data_empi_acc_f_phq$Provider)

##########################
### some preprocessing ###
##########################

#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
data_empi_acc_f_phq[27:29] <- apply(X = data_empi_acc_f_phq[27:29], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq[31:45] <- apply(X = data_empi_acc_f_phq[31:45], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)

########################################
#### Start counting / making subsets ###
########################################

# number of unique scanners and scans
print(paste0("Number of unique procedure names (ie. MRI Head w/out contrast; MRI Source Images): ",  length(unique(data_empi_acc_f_phq$PROC_CODE)))) #n = 11
print(paste0("Number of unique department names (ie. RAD HUP PCAM Ground): ",  length(unique(data_empi_acc_f_phq$DEPARTMENT_NAME)))) #n = 3
print(paste0("Number of unique modalities (ie. HUP PCAM GR MR4 3T; MR6 3T): ",  length(unique(data_empi_acc_f_phq$MODALITY)))) #n = 11
print(paste0("Number of unique accession #s n= ", length(unique(data_empi_acc_f_phq$ACCESSION_NUM)))) # n = 17067

#### PHQ-2 and PHQ-9 summaries
#phq-2;
print(paste0("N PHQ-2 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.2)))) #n = 4701

#phq-9
print(paste0("N PHQ-9 = ", sum(!is.na(data_empi_acc_f_phq$PHQ.9)))) #n = 274

#both phq2 and 9

num_both_phq2_and_9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & !is.na(data_empi_acc_f_phq$PHQ.9)))
print(paste0("num people with both phq2 and 9 = ", num_both_phq2_and_9)) #n = 0

#looking at just people with phq-2 and 9
phq2_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.2)),] #n=4701
phq9_subset <- data_empi_acc_f_phq[which(!is.na(data_empi_acc_f_phq$PHQ.9)),] #n=274

#Number of unique EMPIs in phq2/phq9 subsets
unique_phq2_empis <- length(unique(phq2_subset$EMPI)) #n=966
unique_phq9_empis <- length(unique(phq9_subset$EMPI)) #n=76

#### ICD 9/10 for Depression
dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2339
num_depression <- dim(dep_df)[1]
unique_empi <- which(unique(data_empi_acc_f_phq$EMPI) %in% data_empi_acc_f_phq$EMPI) #index of the locations of unique empis; n = 3885
num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 633

print(paste0("N with ICD 9/10 codes for depression = ", num_depression))

print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep, " out of ", dim(data_empi_acc_f_phq)[1]))

##### PHQ2 or 9 + ICD 9/10 Depression
num_phq2_and_dep <- length(which(!is.na(dep_df$PHQ.2))) #n = 1049
print(paste0("N with ICD 9/10 codes for depression and phq2 (ALL SCANS) = ", num_phq2_and_dep, " out of ", dim(dep_df)[1]))

num_phq9_and_dep <- length(which(!is.na(dep_df$PHQ.9))) #n=153
print(paste0("N with ICD 9/10 codes for depression and phq9 (ALL SCANS) = ", num_phq9_and_dep, " out of ", dim(dep_df)[1]))

#### # unique people with dep dx and phq2/9 -> working on this
dep_dx_phq2_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.2)])) #n = 266
print(paste0("N with ICD 9/10 codes for depression and phq2 (unique) = ", dep_dx_phq2_unique, " out of ", dim(dep_df)[1]))

dep_dx_phq9_unique <- length(unique(dep_df$EMPI[!is.na(dep_df$PHQ.9)])) #n = 51
print(paste0("N with ICD 9/10 codes for depression and phq9 (unique) = ", dep_dx_phq9_unique, " out of ", dim(dep_df)[1]))

####

#demographics and lab historgrams
#print histogram of phq-2/9s for people

hist(data_empi_acc_f_phq$PHQ.2[which(!is.na(data_empi_acc_f_phq$PHQ.2))], main = paste0("PHQ-2 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))), xlab = "PHQ-2", breaks = 8)

hist(data_empi_acc_f_phq$PHQ.9[which(!is.na(data_empi_acc_f_phq$PHQ.9))], main = paste0("PHQ-9 Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$PHQ.9)))), xlab = "PHQ-9", breaks = 30)

#fcode histos
hist(dep_df$PHQ.2, main = paste0("PHQ2 and Dep Histo : n = ", num_phq2_and_dep), xlab = paste0("PHQ2"))

hist(dep_df$PHQ.9,main = paste0("PHQ9 and Dep Histo : n = ", num_phq9_and_dep), xlab = paste0("PHQ9"))

###

#age histogram
hist(data_empi_acc_f_phq$MRI_ENC_AGE[which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE))], main = paste0("Age Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$MRI_ENC_AGE)))), xlab = "AGE AT MRI ENCOUNTER")

#tsh
hist(data_empi_acc_f_phq$TSH[which(!is.na(data_empi_acc_f_phq$TSH))], main = paste0("TSH Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$TSH)))), xlab = "TSH")

#hemoglobin
hist(data_empi_acc_f_phq$hemoglobin[which(!is.na(data_empi_acc_f_phq$hemoglobin))], main = paste0("Hemoglobin Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$hemoglobin)))), xlab = "hemoglobin")

#folate
hist(data_empi_acc_f_phq$FOLATE[which(!is.na(data_empi_acc_f_phq$FOLATE))], main = paste0("Folate Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$FOLATE)))), xlab = "folate", breaks = 10)

#vitamin D
hist(data_empi_acc_f_phq$VitD[which(!is.na(data_empi_acc_f_phq$VitD))], main = paste0("Vitamin D Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$FOLATE)))), xlab = "Vitamin D", breaks = 10)

#sex
hist(data_empi_acc_f_phq$sex_binarized[which(!is.na(data_empi_acc_f_phq$sex_binarized))], 
     main = paste0("Sex Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$sex_binarized))), "; M = ", length(which(data_empi_acc_f_phq$sex_binarized == 1)), "; F = ", length(which(data_empi_acc_f_phq$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

#race
hist(data_empi_acc_f_phq$race_binarized[which(!is.na(data_empi_acc_f_phq$race_binarized))], 
     main = paste0("Race Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$race))), "; White = ", length(which(data_empi_acc_f_phq$race_binarized == 1)), "; NW = ", length(which(data_empi_acc_f_phq$race_binarized == 2))),
     xlab = "white = 1, other = 2", 
     breaks = 2)

#f-codes - people usually have BOTH icd9/10 or neither

```