### 2022-08-13 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
graphics

---
title: "MS Depression Paper"
author: "Erica Baller"
date: '2022-07-28'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(boot)
library(boot.pval)
library(lsr)
#functions
source("~/BBL/msdepression/scripts/msdepression_functions.R")
```

## MS Depression

Welcome to the MS Depression Data, installment 1 analysis markdown! This project was designed to ask a relatively simple question - are patients with MS who are depressed more likely to have lesions in fascicles that connect areas previously associated with depression. The literature are sparse in this domain. It is very clear that depression occurs in up to 50% of patients with MS, and some studies have suggested that it relates to lesion burden. But to my knowledge there are no papers that evaluate how lesions disrupt white matter tracts that serve as the backbone for functional networks. Over the course of the next few chunks, I'll walk through our analysis. It combines imaging data from research grade scans acquired at 3T as part of a clinical protocol (T1 and T2 Flair), and clinical data from the electronic medical record.

## R Markdown

Part 1: 
The first chunk here is reading in appropriate data frames and preprocessing. Two big data frames are read in. The first is the demographic, drug, lab info for the MS patients (from the DAC pull). The second is the data frame of fascicle proportion affected for people with good mimosa (# volume affected/total volume of fascicle). The most important preprocessing occurs in a series of mutate commands. These essentially add new columns to the data frame that can later be used. The thrust of these mutate commands is to define who has depression and who does not.

Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion. Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses. For these preliminary analysis, our inclusion/exclusion are more stringent. Of note, the graph below is irrespective of whether they had good/bad MIMoSA. As such, the MIMoSA subset is smaller.

<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>

Inclusion criteria below:

Inclusion MS Criteria (dataframe read in here has already been filtered down using the exclusions below previously)
  1) Scanned under MS protocol (42K)
  2) MS Diagnosis (17K)
  3) Seen by MS doctor (16K)

Inclusion criteria for Depressed Group
  1) Have an ICD10 F3* code (in our sample, people have F32 (depressive episode), F33 (major depressive disorder), and F34 (Persistent mood [affective] disorders)) 
        - there are NO participants with bipolar disorder, manic episode, or unspecified mood/affective disorder)
        - while participants may have been diagnosed with F06.3 (Depressive disorder due to known physiological condition, with depressive features), they must ALSO have had an F3(2-4) condition
  2) Screened positive for depression on PHQ2 or PHQ9 at any point
  3) Prescribed antidepressant medications per NAMI website
        - https://www.nami.org/About-Mental-Illness/Treatments/Mental-Health-Medications

Inclusion criteria for Healthy Group
  1) No F* diagnoses
  2) At least 1 PHQ2 or PHQ9 with a documented 0
  3) Free of psychiatric medications per NAMI website

Exclusions
  If you received the MS protocol or had a diagnosis of MS but never actually saw an MS provider, we cannot be sure that the diagnosis of MS is truly valid, as it is hard to diagnose anyway.
  If you did not have a diagnosis of depression and never completed a PHQ2 or 9, and had no history of medications, we could not safely confirm nor deny the presence of depression. These subjects were excluded from future analyses. We could consider adding these to the healthy group maybe, but not too inclined to do that right now.

Imaging exclusion
  We are currently only using people who had good MIMOSA (75 or 100). That severely limits our numbers, but that is okay! When we re-QC, we will probably get a lot more into this group.

Final N demographics 
  Depressed: 859 (232 unique)
  Super clean healthy: 604 (148 unique)

(unique, with good MIMoSA):

                         Healthy          Depressed             p      test
  n                                148            232                     
  Race (%)         Caucasian       108 ( 73.0)    173 ( 74.6)   0.821     
                   Non-caucasian    40 ( 27.0)     59 ( 25.4)             
  Sex (%)          Female          117 ( 79.1)    199 ( 85.8)   0.117     
                   Male             31 ( 20.9)     33 ( 14.2)             
  Age (mean (SD))                46.57 (12.66)  48.75 (11.75)   0.089     
  Depression (%)   1               148 (100.0)      0 (  0.0)  <0.001     
                   2                 0 (  0.0)    232 (100.0)             
  PHQ2 (mean (SD))                0.00 (0.00)    0.58 (1.53)   <0.001     
  PHQ9 (mean (SD))                0.00 (NA)     10.43 (9.02)       NA

Networks

  A note on the "depression network." This network was made by Shan Siddiqi in Michael Fox's group by doing a conglomeration of TMS datasets, stroke datasets, Nature Human Behavior 2021 (https://www.nature.com/articles/s41562-021-01161-1). It is an unthresholded mask (he initially sent one where the unit is r, but then sent us one where unit is T). 
  Per Siddiqi's article: The location of each lesion or brain stimulation site (Fig. 2a–c, top panels) was mapped to an underlying brain circuit using a large normative connectome database (n=1000) and previously validated methods (Fig. 2a–c, bottom panels)5. The normative connectome was used to estimate connectivity of each lesion or stimulation site to every voxel in the brain. At each voxel, a Pearson r-value was computed for the correlation between depression score and lesion or stimulation site connectivity to that voxel (Fig. 2a–c, right panels), yielding a population-derived “circuit map” for each of the 14 datasets (Supplementary Figure 1). We generated a combined depression circuit map by taking the mean of all 14 circuit maps, weighted by the sample size of each dataset (Fig. 5a). Of note, lesions are expected to positively correlate with symptoms, whereas the inverse of tms sites and was (so connectivity in these areas was inverted)

  To use it, I thresholded the clusters (3.09), binarized it and then used it as an ROI and calculated, per fascicle, the proportion of the volume occupied by the fascicle that intersected with the depression mask to the whole volume of the fascicle. Most fascicles were either entirely overlapping or entirely outside the depression mask, though the range of overlap was anywhere from 1-30%. This required us to make some choices. We considered an option where the "depression mask" was anywhere that ever overlapped, and non-depression were areas where there was no overlap. We also looked at quartiles, and settled on the top 25% (top quartile), i.e. the top 25% of fascicles with the highest proportion of network overlap were considered in the depression network. Everything outside of that was considered "non_depression" network, or nondep_net. Cranial nerves are too small to reliably be assess with our method, and so they were removed from the analysis.In total, 77 fascicles were evaluated.

Depression top quartile, r&l separated
 [1] "SLF3_R"  "AF_R"    "AF_L"    "SLF2_R"  "CBT_R"   "FAT_R"   "CS_P_R"  "C_PHP_R" "MdLF_L"  "FAT_L"   "SLF2_L"  "MdLF_R"  "CBT_L"   "TR_P_L"  "TR_S_R" 
[16] "CPT_O_R" "TR_P_R"  "PAT_L"   "PAT_R"   "EMC_R"   "TR_S_L"  "CC"

Notebook Overview

Analyses
  There are two main types of analyses I performed, t.tests (for example, to compare inside versus outside depression network) and linear models (for example, looking at the relationship of depression to each individual fascicle). With respect to T-tests, I assume unequal variance for all. For linear models, I look only at the first scan. The general pattern is to look at the total proportion of volume lost across fascicles, then within depression network, and within non depression network. Lastly, I look at overall mean lesion burden irrespective of location by counting the number of nonzero voxels in each individual's mimosa mask.

Results
  Depression group: 232 with depression, 148 healthy:

  A) Lesion volume between depressed and controls

    - P=0.067, NS

  B) Depressed vs Controls Fascicle Affected - Proportions

    - Dep vs non-dep, total fascicle proportion affected p = 0.050 (NS), Dep = 31%, Non dep = 27%
    - Dep vs non, top 25%, **p = 0.047**, Dep = 32%, non dep = 26% 
    - Dep vs non, bottom 75%, no p = 0.057, dep = 30%, non dep = 27%

  C) Depressed vs Controls Fascicle Affected - Volume

    - Dep vs nondep, total fascicle volume affected: **p = 0.045**, Dep = 9382, Non dep = 8260
    - Dep vs nondep in top 25% yes **p = 0.034**, Dep = 4829, non dep = 3914
    - Dep vs nondep in bottom 75%, no p = 0.054, dep = 8826, non dep = 7886

  D) Linear models of values in fascicle associated with Overlap in Depression map

    - T values correlated w/proportion of overlap of fascicle in depression map
        - p of lm (0.010),
        - 95%   ( 0.701,  4.553 )
        - boostrap p calc from CI - **0.0087**

    - Cohen'd correlated w/proportion of overlap of fascicle w/depression map
        - p of lm (0.001)
        -

    - T values correlated w/volume of overlap with depression network (i.e. # voxels that fascicle shares w/depression map),
        - p = 0.031
        - 95%   ( 1.052,  3.724 )
        - bootstrap p calc from CI- **1*10^-16**

    - Cohen'd correlated w/volume of overlap of fascicle w/depression map
        - p of lm (p < 2.2 * 10^-16)
        - Largest effect in corpus callosum
I think it works best when sticking with *volumes* rather than proportions

#### Promis data, late edition

We decided to look at some of the promis data (which measures broad functional categories, though not specifically psychiatric diagnosis) to see if a) measures differed by diagnostic groups and b) whether there were dimensional effects. Unfortunately, the PROMIS picked up in 2018, which was also the same year that Melissa ended her pull, so our n's are small.

Current ns for people with at least 1 promis score= 
Total n promis scores: 1694 
Total n unique individuals w/promis scores: 501

Different scores have different # of overall scans
                        Total                      General_Health                  Quality_of_Life                  Physical_Health 
                            1694                             1693                             1503                             1503 
          Mental_Health_and_Mood   Social_Activities_Satisfaction   Carrying_Out_Social_Activities Carrying_Out_Physical_Activities 
                            1503                             1503                             1501                             1499 
              Emotional_Problems                  Fatique_Average     PostOp.PROMIS.Physical.Score       PostOp.PROMIS.Mental.Score 
                            1499                             1499                             1440                             1442

W/good mimosa:
Total n promis scores : 398
Total n unique individuals w/promis scores: 116
Total n unique indivs in nondepressed group = 36
Total n unique individuals in depressed group = 46

Method:
To use the promis scores, I cleaned them in a separate script (promis_score_cleaning_2010-2022.Rmd)
In this script, I link each scan with its most proximal promis score. The challenge here is that there is no time linking. Summary of the time (in days and years) between scanning and a promis (these are absolute values, do not indicate before or after).

Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    1.0   376.8   763.0   873.3  1229.8  2487.0 
    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
0.002747 1.035027 2.096154 2.399132 3.378434 6.832418

Next, I looked at overall differences between diagnostic groups on each of the PROMIS measures, correcting for multiple comparisons w/FDR (Q<0.05).

Then, I looked at the relationship between promis scores and fascicle measures (overall, 1620 analyses done):

Proportion:
Proportion within versus outside depression map correlated w/promis scores (33 analyses per group (11 promis domains x 3 networks [total, depression 25%, nondep 75%]))
    - all participants (116)
    - Depressed + nondepressed (82)
    - Non-depressed alone (36)
    - Depressed alone (46)

Proportion of individual fascicle affected related to promis scores (777 analyses per group (11 promis domains x 77 individual fascicles))
  - all participants (116)
  - Depressed + nondepressed (82)
  - Non-depressed alone (36)
  - Depressed alone (46)

Volume:

Volume within versus outside depression map correlated w/promis scores (33 analyses per group (11 promis domains x 3 networks [total, depression 25%, nondep 75%]))
   - all participants (116)
    - Depressed + nondepressed (82)
    - Non-depressed alone (36)
    - Depressed alone (46)

Volume of individual fascicle affected related to promis scores  (777 analyses per group (11 promis domains x 77 individual fascicles))
  - all participants (116)
  - Depressed + nondepressed (82)
  - Non-depressed alone (36)
  - Depressed alone (46)

Promis results:
Interestingly, across all 11 scores, with the exception of fatigue, Depressed patients have lower scores (lower = worse functioning), than non-depressed patients. controls. They all fdr correct. What I love about this is fatigue not being different. If fatigue was worse in depressed patients, one could argue that quality of life, health, emotional problems, social activities, etc are all worse because of underlying fatigue. Similarly, as fatigue is a component of the PHQ9 and depression screening, this could be confounding. That they were different in a lot of measures but not fatigue is informative.

Quality_of_Life	0.015199144			
Physical_Health	0.012579264			
Mental_Health_and_Mood	0.012579264			
Social_Activities_Satisfaction	0.001063449			
Carrying_Out_Social_Activities	0.011375799			
Carrying_Out_Physical_Activities	0.031616473			
Emotional_Problems	0.011270683			
Fatique_Average	0.231497853			
PostOp.PROMIS.Physical.Score	0.041172446			
PostOp.PROMIS.Mental.Score	0.001253404

Proportion
Proportion within versus outside depression map correlated w/promis scores - NO SIGNIFICANT FINDINGS
    - all participants (116)
    - Depressed + nondepressed (82)
    - Non-depressed alone (36)
    - Depressed alone (46)

Proportion of individual fascicle affected related to promis scores-  NO SIGNIFICANT FINDINGS
  - all participants (116)
  - Depressed + nondepressed (82)
  - Non-depressed alone (36)
  - Depressed alone (46)

Volume     
Volume within versus outside depression map correlated w/promis scores -  NO SIGNIFICANT FINDINGS
   - all participants (116)
    - Depressed + nondepressed (82)
    - Non-depressed alone (36)
    - Depressed alone (46)

Volume of individual fascicle affected related to promis scores-  NO SIGNIFICANT FINDINGS
  - all participants (116)
  - Depressed + nondepressed (82)
  - Non-depressed alone (36)
  - Depressed alone (46)

#### Using volume rather than proportion to determine inside vs outside depression network
It occurred to me that defining the top 25% depression network has been done by looking at proportion overlap, where everything else seems to work best in volume. It made me wonder what would happen if I defined my top 25% network based on the absolute volume of overlap a fascicle had with the depression network, and ranked based on that (rather than the % of fascicle that overlapped). It tests a slightly different hypothesis, where the depression network is though of as a collection of fascicles that have the most dead ending or overlap, i.e. contribute the most volume to the network. Proportion, on the other hand, thinks more about the specificity of the fascicle - i.e. if the majority of a fascicle dead ends in the network, its activity is probably highly specific to the network. In a volume only, the value of that fascicle's contribution would be neglibile. Using volume = house of representatives. using overlap - senate.

Or put another way, if you do by volume, big fascicles get the most potential votes, by virtue of their bigness. It would be entirely possible for the AF, for example, to have some of the highest overlap with respect to volume, but if that volume is only 10% of the AF fascicle, it means that 90% dead end outside of the depression network. If a tiny fascicle is entirely hit, at least we know there is some specificity to its existence. Conversely, if the AF accounts for 99% of the volume overlap w/the depression network, even if it only 10% of the volume of the AF, it seems like it should be valued highly as part of the depression network. Luckily, results from both work out well.

Results after defining the depresion 25% network based on the volume rather than proportion overlap: i.e. a fascicle is in the "depression network" if the volume that it shares with the depression network is in the top 25%. 
  - Irrespective of network, overall disease by volume: Dep > nondep (p=0.045). 
  - Within dep network: dep disease > nondep disease i, p=0.0.034,
  - Outside dep network (bottom 75%), dep disease ~= nondep disease, p = 0.071 (NS)

dep_network_names_top_quartile_no_cranial_nerves_by_vol
 [1] CC      AF_L    SLF2_R  SLF3_R  FAT_L   TR_S_R  SLF2_L  AF_R    TR_S_L  FAT_R   CS_S_R  CPT_F_R CPT_P_L ILF_R   CPT_P_R CS_S_L  PAT_R   MdLF_L  CBT_R

dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol
  [1] C_FPH_L C_FPH_R C_FP_L  C_FP_R  C_PH_L  C_PHP_L C_PHP_R C_PH_R  C_R_L   C_R_R   EMC_L   EMC_R   IFOF_L  IFOF_R  ILF_L   MdLF_R  PAT_L   SLF1_L  SLF1_R 
[20] SLF3_L  UF_L    UF_R    VOF_L   VOF_R   CB_L    CB_R    ICP_L   ICP_R   MCP     SCP     V       AR_L    AR_R    CBT_L   CPT_F_L CPT_O_L CPT_O_R CS_A_L 
[39] CS_A_R  CS_P_L  CS_P_R  CST_L   CST_R   DRTT_L  DRTT_R  F_L     F_R     ML_L    ML_R    OR_L    OR_R    RST_L   RST_R   TR_A_L  TR_A_R  TR_P_L  TR_P_R 
[58] AC

```{r read_in_df_and_subset}
homedir <- "/Users/eballer/BBL/msdepression/"

data <- read.csv(paste0(homedir, "/data/drugs_data/parsable_msdepression.csv"), sep = ",", header = TRUE) #n=16830, only has people with MS ICD10 code G35 and seen by MS provider

columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE","hemoglobin", "WBC","RBC","B12","FOLATE", "TSH", "RPR","CSFAPPEAR1","CSFAPPEAR4","CSFCOLOR1",       "CSFCOLOR4","CSFTUBE","CSFTUBE4","VitD","PHQ.2","PHQ.9")

##########################
### some preprocessing ###
##########################
#we need to keep scans from people seen by an MS provider (n = 17067) who have an MS diagnostic code (n=16,830)
#we make a bunch of columns from character to integer, binarize sex and race, and put date into a nice format YYYYMMDD, the %m%d%y indicates what it started out as

#goal is to keep people who were seen by 
 data_empi_acc_f_phq <- data %>% 
   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = "")) %>%
   mutate(BEGIN_EXAM_DTTM = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% #make sure to actually save it as date object for later processing
   mutate(EMPI = as.factor(EMPI)) %>%
   mutate(On.Psych.Meds = ifelse(On.Psych.Meds == "True", 1, 0)) %>%
   mutate(On.Antidepressants = ifelse(On.Antidepressants == "True", 1, 0)) %>%
   mutate(Has.PHQ2 = ifelse(Has.PHQ2 == "True", 1, 0)) %>%
   mutate(Has.PHQ9 = ifelse(Has.PHQ9 == "True", 1, 0)) %>%
   mutate(Has.depdx = ifelse(grepl("F3", ICD10), 1, 0)) %>%
   mutate(PHQ.2_modsev_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 >= 3), 1, 0)) %>%
   mutate(PHQ.9_modsev_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 >= 10), 1, 0)) %>%
   mutate(PHQ.2_mild_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 > 0 & PHQ.2 < 3), 1, 0)) %>%
   mutate(PHQ.9_mild_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 > 0 & PHQ.9 < 10), 1, 0)) %>%
   mutate(PHQ.2_zero = ifelse((!is.na(PHQ.2) & PHQ.2 == 0), 1, 0)) %>%
   mutate(PHQ.9_zero = ifelse((!is.na(PHQ.9) & PHQ.9 == 0), 1, 0)) %>%
   mutate(dep_by_dx_phq = ifelse((Has.depdx | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(dep_by_dx_phq_antidep = ifelse((Has.depdx | On.Antidepressants | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(true_healthy = ifelse((!Has.depdx & !On.Psych.Meds & (PHQ.2_zero | PHQ.9_zero)), 1, 0)) %>%
   mutate(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds = ifelse(dep_by_dx_phq_antidep | true_healthy, 1, 0)) %>%
   rowwise(ACCESSION_NUM) %>%
   mutate(depGroupVar =
            sum(c(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds,dep_by_dx_phq_antidep))) %>%#you get an extra point if you are in the depression group, so the end result is dep = 2, healthy = 1, 0 for exclude
   ungroup()

##########################################################
#             Read in Fascicle info             #
##########################################################

fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77
#write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)

########
#fascicle volumes in depression network
########

fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")

#add column to indicate whether does not meet criteria for inclusion in dep mask (0), any overlapping voxels (1), or >10% overlapping voxels (2)
fascicle_volumes_dep_network <- fascicle_volumes_dep %>%
  replace(is.na(.), 0) %>%
  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))

#Write out the new network groupings to a file, will be used later
fascicle_names_dep_network_grouping <- paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping_3_09.csv")
write.table(fascicle_volumes_dep_network, file=fascicle_names_dep_network_grouping, sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)

#fascicle_mapping, returns numerical mapping as well as names of tracts, and whether any overlap with depression network (anything above 1 is overlap), as well as whether the fascicle overlaps, and the proportion of overlap with yeo networks
fascicle_bundle_mapping <- get_fascicle_bundle_mapping_generic_dep_mask(dep_network_grouping_file = fascicle_names_dep_network_grouping)

##########################################################
#             Read in Lesion Volumes                     #
##########################################################

# contains the volume of lesioned voxels, i.e. a sum of all the areas of each individual's binary mask that are 1

lesion_volumes <- read.csv(paste0(homedir, "results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes"), sep = ",", header = TRUE)

##########################################################
#             Merge with data_empi              #
##########################################################
#merge whole data set with fascicle proportions
df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962

#merge with the df that has lesion volumes, 1 per perosn
df_demo_and_fascicles <- merge(df_demo_and_fascicles, lesion_volumes, by = c("EMPI", "EXAM_DATE")) #n = 2962

##########################################################
#            Read in PROMIS Data              #
##########################################################
promis_data <- readRDS(paste0(homedir, "/data/dac/promis_data_cleaned_n1694_unique_indiv_n501_2015-2022.rds"))
promis_score_names <- names(promis_data[7:16])

df_people_with_promis_data <- df_demo_and_fascicles %>%
  filter(EMPI %in% promis_data$EMPI) #n=398

unique_empis_in_promis_group <- unique(df_people_with_promis_data$EMPI) #n=116

```
```{r make_df_with_volume_instead_of_proportion}
#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected

#read in fiber volume values
fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)

#extract data frame that only includes the fascicle proportion lost
df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]

#multiply each row by the vector of total volumes
df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
                  STATS = fiber_volume_values$volume_full, FUN = "*")

#combine back with data_fram
df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)

```

```{r promis_data_handling}
#go through each person w/promis data and append the columns to their row that reflects the closest collected PROMIS

#function, take an empi and date, returns a column of promis scores
get_closest_promis_to_ms_exam_date <- function(subj_empi, subj_exam_date) {
  df_closest_promis <- promis_data %>%
    filter(EMPI == subj_empi) %>%
    mutate(distance_from_scan = abs(RECORDED_TIME-subj_exam_date)) %>%
    filter(abs(RECORDED_TIME-subj_exam_date)==min(abs(RECORDED_TIME-subj_exam_date))) %>%
    slice(1)
  return(df_closest_promis)
}

#make dataframe to store promis info
promis_info_closest_proximity <- data.frame(matrix(NA, nrow=dim(df_people_with_promis_data)[1], ncol=(dim(promis_data)[2] + 1)))
names(promis_info_closest_proximity) <- c(names(promis_data), "distance_from_scan")

for (line in 1:dim(df_people_with_promis_data)[1]) {
  #pull out subject empi and exam date
  print(line)
  subj_empi <- df_people_with_promis_data$EMPI[line]
  subj_exam_date <- df_people_with_promis_data$BEGIN_EXAM_DTTM[line]
  closest_promis_data <- get_closest_promis_to_ms_exam_date(subj_empi = subj_empi, subj_exam_date = subj_exam_date)
  promis_info_closest_proximity[line,] = closest_promis_data
}

#drop unnecessary columns c("DE_PAT_ID", "EMPI", "HUP_MRN", "SEX", "RACE")
promis_info_closest_proximity_just_scores <- promis_info_closest_proximity[,6:dim(promis_info_closest_proximity)[2]]

#combine with df_people_with_promis_data
df_people_with_promis_data_scores_added = cbind(df_people_with_promis_data, promis_info_closest_proximity_just_scores)

#summarize the distance from scans

summary(df_people_with_promis_data_scores_added$distance_from_scan) #mean 873.3 days, median 763 days, range 1 to 2487 days, convert to years
hist(df_people_with_promis_data_scores_added$distance_from_scan) #convert to years

summary(df_people_with_promis_data_scores_added$distance_from_scan/364) #mean 2.40 ys, median 2.10 years, range 0.003 y to 6.83 years
hist(df_people_with_promis_data_scores_added$distance_from_scan/364) #convert to years

#sub volume for proportion
#extract data frame that only includes the fascicle proportion lost
df_fascicles_only <- df_people_with_promis_data_scores_added[ , (names(df_people_with_promis_data_scores_added) %in% fascicle_names)]
df_everything_else <- df_people_with_promis_data_scores_added[ , !(names(df_people_with_promis_data_scores_added) %in% fascicle_names)]

#multiply each row by the vector of total volumes
df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
                  STATS = fiber_volume_values$volume_full, FUN = "*")

#combine back with data_fram
df_demo_and_fascicle_volumes_full_w_promis <- cbind(df_everything_else, df_fascicle_volumes)

```

```{r demographics}

#this will display a demographics table comparing age, race, sex between healthy and depressed groups, using the first presentation to treatment

make_demographics_table_ms(data_frame = data_empi_acc_f_phq)

df_unique_empi <- data_empi_acc_f_phq %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

make_demographics_table_ms(data_frame = df_unique_empi)

df_just_healthy_and_depressed <- df_unique_empi %>% filter(depGroupVar != 0)

make_demographics_table_ms(data_frame = df_just_healthy_and_depressed)

#just people with good mimosa
df_good_mimosa <- df_demo_and_fascicles %>% filter(depGroupVar != 0)
make_demographics_table_ms(data_frame = df_good_mimosa)

df_good_mimosa_unique <- df_demo_and_fascicles %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique)

df_good_mimosa_unique_dep_and_healthy <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique_dep_and_healthy)

### with promis data
make_demographics_table_ms_w_promis(data_frame = df_people_with_promis_data_scores_added)

## overall
#just good mimosa w/promis
df_good_mimosa_promis <- df_people_with_promis_data_scores_added %>% filter(depGroupVar != 0)
make_demographics_table_ms_w_promis(df_good_mimosa_promis) #n=287 scans

#unique
df_people_with_promis_data_scores_added_unique <- df_people_with_promis_data_scores_added %>%
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

make_demographics_table_ms_w_promis(data_frame = df_people_with_promis_data_scores_added_unique)

```

```{r compare_overall_volume_of_lesions_dep_vs_non_dep}

#separate out only people in depression group, and take first instance
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

dep_vs_healthy_t_test_lesion_volume <-  t.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #p=0.07

Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped
barplot_lesion_volume <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ggtitle(paste0("Volume of Mimosa Lesions by Diagnosis, p=", round(dep_vs_healthy_t_test_lesion_volume$p.value,3))) +
  ylab("Volume of Lesion") +
  theme(title = element_text(size=16),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        aspect.ratio = 5/5,
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
#  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
print(barplot_lesion_volume) #p=0.067

barplot_lesion_volume_range <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ggtitle(paste0("Volume of Mimosa Lesions by Diagnosis, p=", round(dep_vs_healthy_t_test_lesion_volume$p.value,3))) +
  ylab("Volume of Lesion") +
  ylim(0,20000) +
  theme(title = element_text(size=16, hjust = 0.2),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        aspect.ratio = 5/5,
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
 # scale_y_continuous(expand = c(0, 0), limits = c(0, 20000)) #comment out if you want colored graph
print(barplot_lesion_volume_range) #p=0.067

##### violin plot

#extract overall volume and diagnosis
df_lesion_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
  dplyr::select(volume_of_mimosa_lesions,depGroupVar) %>%
  filter(depGroupVar != 0) %>% 
  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))

violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, color=depGroupVar)) +#, fill=depGroupVar)) + 
  geom_violin() + #draw_quantiles = c(0.25, 0.5, 0.75)) + 
  stat_summary(fun= "mean", geom = "pointrange", colour = "black") +
  geom_boxplot(width=0.1) + theme_minimal() + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
  ylab("Volume of Lesions") +
  theme(plot.title = element_text(size=16, hjust = 0.8),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
print(violin_lesion_volume) #p=0.067

jitter_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, color=depGroupVar)) +#, fill=depGroupVar)) + 
  geom_jitter(shape=16, position=position_jitter(0.15)) + 
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
  ylab("Volume of Lesions") +
  theme(plot.title = element_text(size=16, hjust = 0.8),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
print(jitter_lesion_volume) #p=0.067

violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, fill=depGroupVar)) +#, fill=depGroupVar)) + 
  geom_violin() + 
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
  ylab("Volume of Lesions") +
  theme(plot.title = element_text(size=16, hjust = 0.8),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
print(violin_lesion_volume) #p=0.067

```

```{r visualize_depression_network_proportion}
##########################################################
#                 Histos                        #
##########################################################

######## Histograms for depression network
#sort based on overlap
fascicle_values_dep_sorted_by_overlap <- fascicle_volumes_dep
fascicle_values_dep_sorted_by_overlap$prop_in_mask[is.na(fascicle_values_dep_sorted_by_overlap$prop_in_mask)] = 0
fascicle_values_dep_sorted_by_overlap$fascicle <- factor(fascicle_values_dep_sorted_by_overlap$fascicle, levels = fascicle_values_dep_sorted_by_overlap$fascicle[order(fascicle_values_dep_sorted_by_overlap$prop_in_mask)])

#remove fascicles with cranial nerves
fascicles_r_and_l_sep_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

plot_all_fascicles_r_and_l_sep_no_cranial_nerves <-ggplot(data = fascicles_r_and_l_sep_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask))+
  ggtitle("Proportion of Volume Overlap per Fascicle 3_09") + 
  geom_bar(stat="identity") +
  xlab("Fascicle") +
  ylab("Overlap with Depression Mask") +
  ylim(0,0.4) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(plot_all_fascicles_r_and_l_sep_no_cranial_nerves)

### remove all rows with cranial nerves in them
# of note, we do not have all cranial nerves represented, we only have, CNII (optic), III(oculomotor), V (trigeminal), VII(facial), VIII(vestibulocochlear)
fascicle_values_dep_no_cranial_nerves = fascicle_values_dep_sorted_by_overlap[(grep(pattern = "CN", x = fascicle_values_dep_sorted_by_overlap$fascicle, invert = TRUE)),]

#figure out how many fascicles that is, then 
quartile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 4),0)

#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take the top 1-quartile
dep_network_names_top_quartile_no_cranial_nerves<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$prop_in_mask, decreasing = TRUE),])$fascicle[1:quartile] #n=19

#get a network that includes all of the fascicles NOT in dep 25% network
dep_network_names_bottom_3_quartiles_no_cranial_nerves <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves)] #n=68

#plot values within the depression network, and values outside of it

fascicles_dep_25_percent_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves),]

plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves <-ggplot(data = fascicles_dep_25_percent_for_plot_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask)) + 
  geom_bar(stat="identity") +
  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, top quartile") + 
  ylim(0,0.4) +
  ylab("Proportion in Depression Mask") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves)

#bottom 3 quartiles
fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves),]

plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves, aes(x = fascicle, y=prop_in_mask)) + 
  geom_bar(stat="identity") +
  ylab("Proportion in Depression Mask") +
  ggtitle("Proportion of Volume per Fascicle Affected Dep net 3_09, bottom three quartiles") + 
  ylim(0,0.4) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5))
print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves) #n=58, which is total 68 - 10 CNs

```

```{r visualize_depression_network_volume}
##########################################################
#                 Histos                        #
##########################################################

######## Histograms for depression network
#sort based on overlap
fascicle_values_dep_sorted_by_overlap <- fascicle_volumes_dep
fascicle_values_dep_sorted_by_overlap$non_zero_voxels_in_dep_map[is.na(fascicle_values_dep_sorted_by_overlap$non_zero_voxels_in_dep_map)] = 0
fascicle_values_dep_sorted_by_overlap$fascicle <- factor(fascicle_values_dep_sorted_by_overlap$fascicle, levels = fascicle_values_dep_sorted_by_overlap$fascicle[order(fascicle_values_dep_sorted_by_overlap$non_zero_voxels_in_dep_map)])

#remove fascicles with cranial nerves
fascicles_r_and_l_sep_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

plot_all_fascicles_r_and_l_sep_no_cranial_nerves <-ggplot(data = fascicles_r_and_l_sep_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map))+
  ggtitle("Volume Overlap per Fascicle 3_09") + 
  geom_bar(stat="identity", fill="006300") +
  xlab("Fascicle") +
  ylab("Overlap with Depression Mask") +
  ylim(0,40000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicles_r_and_l_sep_no_cranial_nerves)

### remove all rows with cranial nerves in them
# of note, we do not have all cranial nerves represented, we only have, CNII (optic), III(oculomotor), V (trigeminal), VII(facial), VIII(vestibulocochlear)
fascicle_values_dep_no_cranial_nerves = fascicle_values_dep_sorted_by_overlap[(grep(pattern = "CN", x = fascicle_values_dep_sorted_by_overlap$fascicle, invert = TRUE)),]

#figure out how many fascicles that is, then 
quartile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 4),0)

#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take the top 1-quartile
dep_network_names_top_quartile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile] #n=19

#get a network that includes all of the fascicles NOT in dep 25% network
dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol)] #n=68

#plot values within the depression network, and values outside of it

fascicles_dep_25_percent_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol),]

plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves <-ggplot(data = fascicles_dep_25_percent_for_plot_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map)) + 
  geom_bar(stat="identity", fill="violet") +
  ggtitle("Volume per Fascicle Affected Dep net 3_09, top quartile") + 
  ylab("Volume in Depression Mask") +
  ylim(0,40000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves)

#bottom 3 quartiles
fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves<- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol),]

plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map)) + 
  geom_bar(stat="identity", fill="darkgreen") +
  ylab("Volume in Depression Mask") +
  ylim(0,40000) +
  ggtitle("Volume per Fascicle Affected Dep net 3_09, bottom three quartiles") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves) #n=58, which is total 68 - 10 CNs

```

```{r proof_of_concept}

#this section creates histograms looking at the proportion of volume lost in each fascicle across subjects
#we recapitulate earlier work and show that higher volume loss with our method reflects regions previously known to have higher volume loss
#These findings suggest that the streamline filtering approach could be a reasonable approach for this analysis

data_empi_acc_f_phq_unique <- data_empi_acc_f_phq %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#have to pull out just the pieces I'm interested in
just_dep_and_empi <- subset(data_empi_acc_f_phq, select = c("EMPI", "depGroupVar"))
added_dep_to_fascicles <- unique(merge(just_dep_and_empi, fascicle_proportions, by = c("EMPI")))

melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))

#exclude regions where there was no lesion at all
lesioned <- subset(melted_df, value > 0)

#only include people in healthy or depressed group
lesioned_dx <- subset(lesioned, depGroupVar != 0)

#separate out healthy and depressed
healthy <- subset(lesioned, depGroupVar == 1)
depressed <- subset(lesioned, depGroupVar == 2)

q<-ggplot(lesioned, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r<-ggplot(lesioned_dx, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r_color<-ggplot(lesioned_dx, aes(x=value, color=factor(depGroupVar))) + 
  geom_histogram(color="darkgreen") + 
  facet_wrap(~variable)
x<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
y<-ggplot(healthy, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
z<-ggplot(depressed, aes(x=value)) + geom_histogram() + facet_wrap(~variable)

#print(q)
#print(r)
print(r_color)
#print(x)
#print(y)
#print(z)

###########################################################################################################
##### Looking at distribution of disease across fascicles across group and within healthy and depressed ###
###########################################################################################################

#### look at mean fascicle volume overlap across participants and transpose
# drop exam date and empi, and transpose
fascicles_alone_df <- t(added_dep_to_fascicles[,!(colnames(added_dep_to_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
fascicle_names <- rownames(fascicles_alone_df)

#take row means
fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
names(fascicles_alone_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

write.table(fascicles_alone_rowMeans, file = paste0(homedir, "/results/fascicle_mean_proportion_overlap_nfasc_77.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans

#make this separate in case I want to use this distribution for ranking depressed and healthy networks. 
order_of_mean_values_overall <- fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)]

fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])

######## Histograms/bargraphs
#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Proportion of Volume per Fascicle Affected") + 
  ylab("Proportion") + 
  geom_bar(stat="identity") + 
 # scale_fill_continuous(type="viridis") + 
  scale_fill_gradient(low = "red", high = "yellow") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)

######### 
#############
#### look at mean fascicle overlap across healthy participants and transpose
#############

# drop exam date and empi, and transpose
healthy_group_fascicles <- added_dep_to_fascicles[(added_dep_to_fascicles$depGroupVar == 1),]
fascicles_alone_df <- t(healthy_group_fascicles[,!(colnames(healthy_group_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
fascicle_names <- rownames(fascicles_alone_df)

#take row means
fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
names(fascicles_alone_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

write.table(fascicles_alone_rowMeans, file = paste0(homedir, "/results/fascicle_mean_proportion_overlap_nfasc_77_nondep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans

fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])

######## Histograms/bargraphs

# ordered by mean value overall
fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
sorted_ggplot_healthy_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values NonDepressed") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") +
  #scale_fill_gradient(low = "red", high = "yellow") +
  ylab("Proportion Affected") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot_healthy_ordered)

###########
#### look at mean fascicle volume across depressed participants and transpose
###########
# drop exam date and empi, and transpose
depressed_group_fascicles <- added_dep_to_fascicles[(added_dep_to_fascicles$depGroupVar == 2),]
fascicles_alone_df <- t(depressed_group_fascicles[,!(colnames(depressed_group_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
fascicle_names <- rownames(fascicles_alone_df)

#take row means
fascicles_alone_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df))
names(fascicles_alone_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_rowMeans <- fascicles_alone_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

write.table(fascicles_alone_rowMeans, file = paste0(homedir, "/results/fascicle_mean_proportion_overlap_nfasc_77_dep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_rowMeans

fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])

######## Histograms/bargraphs

# ordered by mean value overall
fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
  ylab("Proportion Affected") +
  geom_bar(stat="identity") + 
  #scale_fill_gradient(low = "red", high = "yellow") +
  scale_fill_continuous(type="viridis") + #, color="orange") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot_depressed_ordered)

### color by quartile
#num_buckets = round(length(fascicle_values_sorted_by_overlap$fascicle)/4,0) #get the number of quartiles
#color_columns= rep(x = c("#8DD3C7", "#FFFFB3", "#BEBADA", "#FB8072"), times=1, each=num_buckets) #assign each quartile a color
#color_columns= rep(x = c("violet", "blue", "green", "red"), times=1, each=num_buckets) #assign each quartile a color

#if (length(fascicle_values_sorted_by_overlap$fascicle) %% 4 != 0){
 # color_columns=c("#8DD3C7", color_columns) #add one more to the beginning of the vector since the number is odd, so mod is odd number
#}

fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type = "viridis") + 
  ylab("Proportion Affected") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot_depressed_ordered)

##### look at mean fascicles affected in dep vs non dep network
######
#plot values within the depression network, and values outside of it

fascicles_dep_25_percent_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves),]

plot_all_fascicles_r_and_l_sep_top_quartile <-ggplot(data = fascicles_dep_25_percent_for_plot, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Dep net, top quartile") + 
  geom_bar(stat="identity", fill="darkblue") +
  ylab("Proportion") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicles_r_and_l_sep_top_quartile)

#bottom 3 quartiles
fascicles_dep_bottom_3_quartiles_for_plot <- fascicle_values_sorted_by_overlap[(fascicle_values_sorted_by_overlap$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves),]

plot_all_fascicles_r_and_l_sep_bottom_3_quartiles <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected, bottom three quartiles") + 
  ylab("Proportion") +
  geom_bar(stat="identity", fill="red") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles)

############ Look at mean overall volume across all brain regions
#### look at mean fascicle volume overlap across participants and transpose
# drop exam date and empi, and transpose
fascicles_alone_df_vol <- t(df_demo_and_fascicle_volumes_full[,(colnames(df_demo_and_fascicle_volumes_full) %in% fascicle_names_no_cranial_nerves)])
fascicle_names <- rownames(fascicles_alone_df_vol)

#take row means
fascicles_alone_vol_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df_vol))
names(fascicles_alone_vol_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_vol_rowMeans <- fascicles_alone_vol_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

write.table(fascicles_alone_vol_rowMeans, file = paste0(homedir, "/results/fascicle_mean_volume_nfasc_77.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_vol_rowMeans

#make this separate in case I want to use this distribution for ranking depressed and healthy networks. 
order_of_mean_values_overall <- fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)]

fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])

######## Histograms/bargraphs
#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Volume per Fascicle Affected", breaks = 20)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Volume per Fascicle Affected") + 
  ylab("Volume") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") + 
  ylim(0,230000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Volume per Fascicle Affected, exclude CC") + 
  ylab("Volume") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") + 
  ylim(0,30000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)

##### healthy volume
df_demo_and_fascicle_volumes_full_nondep <- df_demo_and_fascicle_volumes_full %>% filter(depGroupVar==1)

# drop exam date and empi, and transpose
fascicles_alone_df_vol <- t(df_demo_and_fascicle_volumes_full_nondep[,(colnames(df_demo_and_fascicle_volumes_full_nondep) %in% fascicle_names_no_cranial_nerves)])
fascicle_names <- rownames(fascicles_alone_df_vol)

#take row means
fascicles_alone_vol_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df_vol))
names(fascicles_alone_vol_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_vol_rowMeans <- fascicles_alone_vol_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

write.table(fascicles_alone_vol_rowMeans, file = paste0(homedir, "/results/fascicle_mean_volume_nfasc_77_nondep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_vol_rowMeans

#make this separate in case I want to use this distribution for ranking depressed and healthy networks. 
order_of_mean_values_overall <- fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)]

fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])

######## Histograms/bargraphs
#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Volume per Fascicle Affected", breaks = 20)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Volume per Fascicle Affected Nondepressed") + 
  ylab("Volume") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") + 
  ylim(0,230000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Volume per Fascicle Affected Nondepressed, exclude CC") + 
  ylab("Volume") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") + 
  ylim(0,30000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)

##### depressed volume
# drop exam date and empi, and transpose

df_demo_and_fascicle_volumes_full_dep <- df_demo_and_fascicle_volumes_full %>% filter(depGroupVar==2)

fascicles_alone_df_vol <- t(df_demo_and_fascicle_volumes_full_dep[,(colnames(df_demo_and_fascicle_volumes_full_dep) %in% fascicle_names_no_cranial_nerves)])
fascicle_names <- rownames(fascicles_alone_df_vol)

#take row means
fascicles_alone_vol_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df_vol))
names(fascicles_alone_vol_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_vol_rowMeans <- fascicles_alone_vol_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

write.table(fascicles_alone_vol_rowMeans, file = paste0(homedir, "/results/fascicle_mean_volume_nfasc_77_dep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_vol_rowMeans

#make this separate in case I want to use this distribution for ranking depressed and healthy networks. 
order_of_mean_values_overall <- fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)]

fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = fascicle_values_sorted_by_overlap$fascicle[order(fascicle_values_sorted_by_overlap$value)])

######## Histograms/bargraphs
#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Volume per Fascicle Affected", breaks = 20)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Volume per Fascicle Affected Depressed") + 
  ylab("Volume") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") + 
  ylim(0,230000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Volume per Fascicle Affected Depressed, exclude CC") + 
  ylab("Volume") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") + 
  ylim(0,30000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)
```

```{r depression_vs_healthy_overall_proportion_of_fascicles_affected}

#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#test whether proportion lost overall different between depressed and healthy- No, p = 0.050, Dep = 31%, Non dep = 27%
total_prop_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054

#look at within depression network using top quartiles, yes p = 0.047, Dep = 32%, non dep = 26%
total_prop_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)

#look at bottom 75%, no p = 0.057, dep = 30%, non dep = 27%
total_prop_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057

############
#bar plot
############

#over all fascicles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)
barplot_overall <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean Proportion of Fascicles Affected") +
  ggtitle(paste0("Total Proportion Affected by dx, p=", round(total_prop_lost_dep_vs_healthy_unique$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph

print(barplot_overall)

#in top quartile

Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_top_quartile <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean proportion of Fascicles Affected") +
    ggtitle(paste0("Total Propotion Affected by dx in Dep 25%, p=", round(total_prop_lost_dep_vs_healthy_unique_in_dep_network_top_quartile$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) 
print(barplot_top_quartile)

#in bottom 3 quartiles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_bottom_3_quartiles <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean Proportion of Fascicles Affected") +
    ggtitle(paste0("Total Proportion Affected by dx, Bottom 75% p=", round(total_prop_lost_dep_vs_healthy_unique_in_bottom_3_quartiles$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52))

print(barplot_bottom_3_quartiles)

```

```{r depression_vs_healthy_overall_volume_of_fascicles_affected}

#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected

#read in fiber volume values
#fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)

#extract data frame that only includes the fascicle proportion lost
#df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
#df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]

#multiply each row by the vector of total volumes
#df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
 #                 STATS = fiber_volume_values$volume_full, FUN = "*")

#combine back with data_fram
#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_vol_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_vol_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#test whether volume lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054

#look at within depression network using top quartiles, yes p = 0.034, Dep = 4829, non dep = 3914
total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)

#look at bottom 75%, no p = 0.054, dep = 8826, non dep = 7886
total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057

############
#bar plot
############

#over all fascicles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)
barplot_overall <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
  ggtitle(paste0("Total Volume Affected by dx, p=", round(total_volume_lost_dep_vs_healthy_unique$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 16),
        aspect.ratio=5/5,
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 20000)) #comment out if you want colored graph

print(barplot_overall)

#in top quartile

Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_top_quartile <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
    ggtitle(paste0("Total Volume Affected by dx in Dep 25%, p=", round(total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 16), 
        aspect.ratio=5/5,
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 20000)) 
print(barplot_top_quartile)

#in bottom 3 quartiles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_bottom_3_quartiles <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
    ggtitle(paste0("Total Volume Affected by dx, Bottom 75% p=", round(total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 16), 
        aspect.ratio=5/5,
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 20000))

print(barplot_bottom_3_quartiles)

######## Violin plots

df_mean_fascicle_volumes_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
dplyr::select(total_fascicle_vol_lost, total_fascicle_vol_lost_indepnet_top_quartile, total_fascicle_vol_lost_nondep_bottom_3_quartiles, depGroupVar) %>%
  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))

jitter_mean_fascicle_volume <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost, color=depGroupVar)) +
  geom_jitter(shape=16, position=position_jitter(0.15)) + 
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Mean Fascicle Volume Affected by Diagnosis")) +
  ylab("Mean Volume of Fascicles") +
  ylim(0,40000) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(jitter_mean_fascicle_volume) #p=0.067

jitter_mean_fascicle_volume_dep <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost_indepnet_top_quartile, color=depGroupVar)) +
  geom_jitter(shape=16, position=position_jitter(0.15)) + 
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Mean Fascicle Volume Affected Dep Net by Diagnosis")) +
  ylab("Mean Volume of Fascicles") +
  ylim(0,40000) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(jitter_mean_fascicle_volume_dep) #p=0.067

jitter_mean_fascicle_volume_nondep <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost_nondep_bottom_3_quartiles, color=depGroupVar)) +
  geom_jitter(shape=16, position=position_jitter(0.15)) + 
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Mean Fascicle Volume Affected Nondep Net by Diagnosis")) +
  ylab("Mean Volume of Fascicles") +
  ylim(0,40000) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(jitter_mean_fascicle_volume_nondep) #p=0.067

##### Geom violin

violin_mean_fascicle_volume <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost, fill=depGroupVar)) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Mean Fascicle Volume Affected by Diagnosis")) +
  ylab("Mean Volume of Fascicles") +
  ylim(0,42000) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(violin_mean_fascicle_volume) #p=0.067

violin_mean_fascicle_volume_dep <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost_indepnet_top_quartile, fill=depGroupVar)) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Mean Fascicle Volume Affected Dep Net by Diagnosis")) +
  ylab("Mean Volume of Fascicles") +
  ylim(0,42000) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(violin_mean_fascicle_volume_dep) #p=0.067

violin_mean_fascicle_volume_nondep <- ggplot(df_mean_fascicle_volumes_and_diagnosis, aes(x=depGroupVar, y=total_fascicle_vol_lost_nondep_bottom_3_quartiles,  fill=depGroupVar)) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Mean Fascicle Volume Affected Nondep Net by Diagnosis")) +
  ylab("Mean Volume of Fascicles") +
  ylim(0,42000) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(violin_mean_fascicle_volume_nondep) 
```

```{r depression_vs_healthy_overall_volume_of_fascicles_affected_dep_net_by_volume}

#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected

#read in fiber volume values
#fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)

#extract data frame that only includes the fascicle proportion lost
#df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
#df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]

#multiply each row by the vector of total volumes
#df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
 #                 STATS = fiber_volume_values$volume_full, FUN = "*")

#combine back with data_fram
#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_vol_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_vol_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves_by_vol))) %>%
  mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#test whether volume lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054

#look at within depression network using top quartiles, yes p = 0.034, Dep = 4829, non dep = 3914
total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)

#look at bottom 75%, no p = 0.054, dep = 8826, non dep = 7886
total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057

############
#bar plot
############

#over all fascicles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)
barplot_overall <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
  ggtitle(paste0("Total Volume Affected by dx, p=", round(total_volume_lost_dep_vs_healthy_unique$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 10000)) #comment out if you want colored graph

print(barplot_overall)

#in top quartile

Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_top_quartile <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
    ggtitle(paste0("Total Volume Affected by dx in Dep 25%, p=", round(total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
  #scale_y_continuous(expand = c(0, 0), limits = c(0, 10000)) 
print(barplot_top_quartile)

#in bottom 3 quartiles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_bottom_3_quartiles <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
    ggtitle(paste0("Total Volume Affected by dx, Bottom 75% p=", round(total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 20000))

print(barplot_bottom_3_quartiles)

```
```{r depression_vs_healthy_overall_volume_of_fascicles_affected_dep_network_fascicles_determined_by_vol}

#goal of this segment is to look at overall lesion burden between depression and controls, outside of fascicles
#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected

#read in fiber volume values
#fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)

#extract data frame that only includes the fascicle proportion lost
#df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
#df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]

#multiply each row by the vector of total volumes
#df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
 #                 STATS = fiber_volume_values$volume_full, FUN = "*")

#combine back with data_fram
#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_vol_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_vol_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_vol_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#test whether volume lost overall different between depressed and healthy- Yes, p = 0.045, Dep = 9382, Non dep = 8260
total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054

#look at within depression network using top quartiles, yes p = 0.034, Dep = 4829, non dep = 3914
total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F)

#look at bottom 75%, no p = 0.054, dep = 8826, non dep = 7886
total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 30%, mean non_dep = 27%, p = 0.057

############
#bar plot
############

#over all fascicles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)
barplot_overall <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
  ggtitle(paste0("Total Volume Affected by dx, p=", round(total_volume_lost_dep_vs_healthy_unique$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 10000)) #comment out if you want colored graph

print(barplot_overall)

#in top quartile

Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_indepnet_top_quartile[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_top_quartile <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
    ggtitle(paste0("Total Volume Affected by dx in Dep 25%, p=", round(total_volume_lost_dep_vs_healthy_unique_in_dep_network_top_quartile$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 10000)) 
print(barplot_top_quartile)

#in bottom 3 quartiles
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_vol_lost_nondep_bottom_3_quartiles[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)

# Grouped

barplot_bottom_3_quartiles <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Mean volume of Fascicles Affected") +
    ggtitle(paste0("Total Volume Affected by dx, Bottom 75% p=", round(total_volume_lost_dep_vs_healthy_unique_in_bottom_3_quartiles$p.value,3))) + 
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        plot.title = element_text(hjust=0.5, size = 14), 
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 10000))

print(barplot_bottom_3_quartiles)

```
```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_prop}
# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups

##############################
## FASCICLES TO BE INCLUDED ##
##############################

#define the depression network (all areas that intersect with depression network, no cranial nerves)
fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77

#######################
### depression map  ###
#######################

### define depression map
fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")

#filter to exclude fascicles w/cranial nerves
fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
  replace(is.na(.), 0) %>%
  filter(fascicle%in% fascicle_names_no_cranial_nerves)

#make data frame that includes each fascicle in 1 column, and each proportion of overlap with depression network in second column

df_fascicle_and_prop_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "prop_in_mask"))

#######################
### real data x dx  ###
#######################

#then, make the data frame
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# lapply to get results of t-test in each fascicle
#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
{

  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
 eval(parse(text=t_test_command))
})
names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves

#pull t-statistic
dep_vs_healthy_t_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$statistic)

#convert to data frame
df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
row.names(df_t_values) <- fascicle_names_no_cranial_nerves

#pull p-values
dep_vs_healthy_p_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$p.value)

#convert to data frame
df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
row.names(df_dep_vs_healthy_p_values) <- fascicle_names_no_cranial_nerves

#FDR correct p-values
dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")

#Convert to data frame
df_dep_vs_healthy_pfdr <- as.data.frame(dep_vs_healthy_pfdr)
row.names(df_dep_vs_healthy_pfdr) <- fascicle_names_no_cranial_nerves

#######################
### plot differences ##
#######################

#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle

df_fascicle_dep_mask_overlap_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_fascicle_and_prop_in_mask$prop_in_mask, df_t_values$dep_vs_healthy_t_values)
names(df_fascicle_dep_mask_overlap_t_value) <- c("fascicle", "depression_mask_overlap", "dep_vs_healthy_t")

#make copy to sort
df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask <- df_fascicle_dep_mask_overlap_t_value

df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask$depression_mask_overlap)])

# bar graph of T-values of fascicles, sorted by depression overlap
plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_prop_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") + 
  geom_bar(stat="identity") +
  ylim(0,5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(plot_fascicle_by_t_sorted_by_dep_overlap)

# plot fascicle overlap x t value
#statistic, p = 0.010
dep_overlap_x_dep_vs_healthy_t <- lm(dep_vs_healthy_t ~ depression_mask_overlap, data=df_fascicle_dep_mask_overlap_t_value)
print(summary(dep_overlap_x_dep_vs_healthy_t))

scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_t_value, aes(x=depression_mask_overlap, y=dep_vs_healthy_t)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Proportion of Fascicle in Depression Network") + 
  ylab("Effect of diagnosis calculated in proportions (T)") + 
  ggtitle(paste0("Fascicle/depression mask Proportion overlap by dx T, p = ", round(summary(dep_overlap_x_dep_vs_healthy_t)$coefficients[2,4],3))) + 
  theme(plot.title=element_text(hjust=0.5))
print(scatter_t_val_by_dep_net_overlap)

```

```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_prop_effect_size}
# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups

##############################
## FASCICLES TO BE INCLUDED ##
##############################

#define the depression network (all areas that intersect with depression network, no cranial nerves)
fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77

#######################
### depression map  ###
#######################

### define depression map
fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")

#filter to exclude fascicles w/cranial nerves
fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
  replace(is.na(.), 0) %>%
  filter(fascicle%in% fascicle_names_no_cranial_nerves)

#make data frame that includes each fascicle in 1 column, and each proportion of overlap with depression network in second column

df_fascicle_and_prop_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "prop_in_mask"))

#######################
### real data x dx  ###
#######################

#then, make the data frame
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# lapply to get results of t-test in each fascicle
#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_no_cranial_nerves, function(x) 
{
  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
  eval(parse(text=text_for_eval))
})
names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_no_cranial_nerves

#convert to data frame
df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)

#######################
### plot differences ##
#######################

#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle

df_fascicle_dep_mask_overlap_d_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_fascicle_and_prop_in_mask$prop_in_mask, df_d_values)
names(df_fascicle_dep_mask_overlap_d_value) <- c("fascicle", "depression_mask_overlap", "dep_vs_healthy_d")

#make copy to sort
df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask <- df_fascicle_dep_mask_overlap_d_value

df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask$depression_mask_overlap)])

# bar graph of T-values of fascicles, sorted by depression overlap
plot_fascicle_by_d_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_d_value_sorted_by_prop_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_d), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") + 
  geom_bar(stat="identity") +
  ylim(-0.1,0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(plot_fascicle_by_d_sorted_by_dep_overlap)

# plot fascicle overlap x t value
#statistic, p = 0.010
dep_overlap_x_dep_vs_healthy_d <- lm(dep_vs_healthy_d ~ depression_mask_overlap, data=df_fascicle_dep_mask_overlap_d_value)
print(summary(dep_overlap_x_dep_vs_healthy_d))

scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_overlap_d_value, aes(x=depression_mask_overlap, y=dep_vs_healthy_d)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Proportion of Fascicle in Depression Network") + 
  ylab("Effect of diagnosis calculated in proportions (D)") + 
  ggtitle(paste0("Fascicle/depression mask Proportion overlap by dx Eff Size, p = ", round(summary(dep_overlap_x_dep_vs_healthy_d)$coefficients[2,4],3))) + 
  theme(plot.title=element_text(hjust=0.5))
print(scatter_d_val_by_dep_net_overlap)

```

```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_volume}

#define the depression network (all areas that intersect with depression network, no cranial nerves)
fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77

#######################
### depression map  ###
#######################

### define depression map
fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")

#filter to exclude fascicles w/cranial nerves
fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
  replace(is.na(.), 0) %>%
  filter(fascicle%in% fascicle_names_no_cranial_nerves)

df_fascicle_and_volume_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))

########################
## Get Volume Values ###
########################

#read in fiber volume values
fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)

#extract data frame that only includes the fascicle proportion lost
df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]

#multiply each row by the vector of total volumes
df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
                  STATS = fiber_volume_values$volume_full, FUN = "*")

#combine back with data_fram
df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)

#isolate out only the depressed group and healthy group, fascicles should be in volumes rather than proportions
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# lapply to get results of t-test in each fascicle
#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_t <- lapply(fascicle_names_no_cranial_nerves, function(x) 
{
  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
  eval(parse(text=t_test_command))
})
names(fascicle_dep_vs_healthy_t) <- fascicle_names_no_cranial_nerves

#pull t-statistic
dep_vs_healthy_t_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$statistic)

#convert to data frame
df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
row.names(df_t_values) <- fascicle_names_no_cranial_nerves

#pull p-values
dep_vs_healthy_p_values <- sapply(fascicle_dep_vs_healthy_t, function(x) x$p.value)

#convert to data frame
df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
row.names(df_dep_vs_healthy_p_values) <- fascicle_names_no_cranial_nerves

#FDR correct p-values
dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")

#Convert to data frame
df_dep_vs_healthy_pfdr <- as.data.frame(dep_vs_healthy_pfdr)
row.names(df_dep_vs_healthy_pfdr) <- fascicle_names_no_cranial_nerves

#######################
### plot differences ##
#######################

#combine data frames, column 1 has names of fascicles, column 2 has proportion of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle

df_fascicle_dep_mask_volume_t_value <- data.frame(df_fascicle_and_prop_in_mask$fascicle, df_t_values$dep_vs_healthy_t_values, df_fascicle_and_volume_in_mask$non_zero_voxels_in_dep_map)
names(df_fascicle_dep_mask_volume_t_value) <- c("fascicle", "dep_vs_healthy_t", "volume_in_dep_mask")

df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask <- df_fascicle_dep_mask_volume_t_value

#order by volume in depression mask
df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$volume_in_dep_mask)])

# bar graph of T-values of fascicles, sorted by depression overlap
plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by volume of dep network") + 
  geom_bar(stat="identity") +
  ylim(0,5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(plot_fascicle_by_t_sorted_by_dep_overlap)

#statistic volume, p = 0.031
dep_overlap_x_dep_vs_healthy_t <- lm(dep_vs_healthy_t ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_t_value)
print(summary(dep_overlap_x_dep_vs_healthy_t))

scatter_t_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_t_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_t)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect of diagnosis (T)") + 
  ggtitle(paste0("Fascicle/depression mask volume by dx T, p = ", round(summary(dep_overlap_x_dep_vs_healthy_t)$coefficients[2,4],3))) + 
  theme(plot.title=element_text(hjust=0.5))
print(scatter_t_val_by_dep_net_overlap)

#remove corpus callosum
df_fascicle_dep_mask_volume_t_value_no_cc <- df_fascicle_dep_mask_volume_t_value %>%
  filter(fascicle != "CC")

df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask <- df_fascicle_dep_mask_volume_t_value_no_cc

#order by volume in depression mask
df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask$volume_in_dep_mask)])

# bar graph of T-values of fascicles, sorted by depression overlap
plot_fascicle_by_t_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_t_value_sorted_by_vol_in_depression_mask, aes(x = fascicle, y=dep_vs_healthy_t), main = "Dep vs Healthy T value, Fascicles ordered by volume of dep network, no CC") + 
  geom_bar(stat="identity") +
  ylim(0,5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(plot_fascicle_by_t_sorted_by_dep_overlap)

#statistic volume, 
dep_overlap_x_dep_vs_healthy_t_no_cc <- lm(dep_vs_healthy_t ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_t_value_no_cc)
print(summary(dep_overlap_x_dep_vs_healthy_t_no_cc))

scatter_t_val_by_dep_net_overlap_no_cc <- ggplot(data=df_fascicle_dep_mask_volume_t_value_no_cc, aes(x=volume_in_dep_mask, y=dep_vs_healthy_t)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect of diagnosis (T)") + 
  ggtitle(paste0("Fascicle/depression mask volume by dx T, no CC, p = ", round(summary(dep_overlap_x_dep_vs_healthy_t_no_cc)$coefficients[2,4],3))) + 
  theme(plot.title=element_text(hjust=0.5))
print(scatter_t_val_by_dep_net_overlap_no_cc)

```

Effect size in volume, w /cc (eff size NOT done in proportions)

```{r comparing_depression_map_overlap_with_dx_diff_no_cranial_nerves_vol_effect_size}
# the purpose of this chunk is to plot the relationship of the degree of overlap of a depression network with the diagnostic difference between groups

##############################
## FASCICLES TO BE INCLUDED ##
##############################

#define the depression network (all areas that intersect with depression network, no cranial nerves)
fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77

#######################
### depression map  ###
#######################

### define depression map
fascicle_volumes_and_props_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")

#filter to exclude fascicles w/cranial nerves
fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_and_props_dep %>%
  replace(is.na(.), 0) %>%
  filter(fascicle%in% fascicle_names_no_cranial_nerves)

df_fascicle_and_vol_in_mask <- subset(x = fascicle_volumes_dep_network_no_cranial_nerves, select = c("fascicle", "non_zero_voxels_in_dep_map"))

########################
## Get Volume Values ###
########################

#read in fiber volume values
fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)

#extract data frame that only includes the fascicle proportion lost
df_fascicles_only <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names)]
df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names)]

#multiply each row by the vector of total volumes
df_fascicle_volumes <- sweep(x = df_fascicles_only, MARGIN = 2,                        # Sweep with Complex STATS
                  STATS = fiber_volume_values$volume_full, FUN = "*")

#combine back with data_fram
df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)

###############################
#######################
### real data x dx  ###
#######################

#then, make the data frame
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# lapply to get results of t-test in each fascicle
#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_no_cranial_nerves, function(x) 
{
  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
  eval(parse(text=text_for_eval))
})
names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_no_cranial_nerves

#convert to data frame
df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)

#######################
### plot differences ##
#######################

#combine data frames, column 1 has names of fascicles, column 2 has Volume of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle

df_fascicle_dep_mask_volume_overlap_d_value <- data.frame(df_fascicle_and_vol_in_mask$fascicle, df_fascicle_and_vol_in_mask$non_zero_voxels_in_dep_map, df_d_values)
names(df_fascicle_dep_mask_volume_overlap_d_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_d")

#make copy to sort
df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask <- df_fascicle_dep_mask_volume_overlap_d_value

df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle  <- factor(df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle, levels = df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$fascicle[order(df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask$volume_in_dep_mask)])

# bar graph of eff size of fascicles, sorted by depression overlap
plot_fascicle_by_d_sorted_by_dep_overlap <-ggplot(data = df_fascicle_dep_mask_overlap_d_value_sorted_by_vol_overlap_depression_mask, aes(x = fascicle, y=dep_vs_healthy_d), main = "Dep vs Healthy T value, Fascicles ordered by overlap with dep network") + 
  geom_bar(stat="identity") +
  #ylim(-0.1,0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
print(plot_fascicle_by_d_sorted_by_dep_overlap)

# plot fascicle overlap x Eff size value
#statistic, p=001
dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value )
print(summary(dep_overlap_x_dep_vs_healthy_d))

scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect Size (D)") + 
  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
#  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network")) + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text = element_text(size=12, color = "black"),
        axis.title = element_text(size=14, color = "black"),
        axis.line = element_line(color="black", size=1.3),
        panel.background = element_blank())
print(scatter_d_val_by_dep_net_overlap)

#no p value
scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect Size (D)") + 
#  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network")) + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text = element_text(size=12, color = "black"),
        axis.title = element_text(size=14, color = "black"),
        axis.line = element_line(color="black", size=1.3),
        panel.background = element_blank())
print(scatter_d_val_by_dep_net_overlap)

#no corpus callosum
df_no_cc <- df_fascicle_dep_mask_volume_overlap_d_value %>% filter(fascicle != "CC")

dep_overlap_x_dep_vs_healthy_vol_d_no_cc <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_no_cc )
print(summary(dep_overlap_x_dep_vs_healthy_d))

scatter_d_val_by_dep_net_overlap_no_cc <- ggplot(data=dep_overlap_x_dep_vs_healthy_vol_d_no_cc, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect Size (D)") + 
 ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d_no_cc)$coefficients[2,4],3))) + 
#  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, excluding CC")) + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text = element_text(size=12, color = "black"),
        axis.title = element_text(size=14, color = "black"),
        axis.line = element_line(color="black", size=1.3),
        panel.background = element_blank())
print(scatter_d_val_by_dep_net_overlap_no_cc)

#no p value
scatter_d_val_by_dep_net_overlap_no_cc <- ggplot(data=dep_overlap_x_dep_vs_healthy_vol_d_no_cc, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect Size (D)") + 
# ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_d_no_cc)$coefficients[2,4],3))) + 
 ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, excluding CC")) + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text = element_text(size=12, color = "black"),
        axis.title = element_text(size=14, color = "black"),
        axis.line = element_line(color="black", size=1.3),
        panel.background = element_blank())
print(scatter_d_val_by_dep_net_overlap_no_cc)

```

```{r bootstrapping_proportion_and_volume_x_dep_t_analyses_and_eff_sizes}

######### functions
tval <- function(formula, data, indices) {
  df <- data[indices,] # allows boot to select sample
  fit <- lm(formula, data=df)
  return(summary(fit)$coefficients[2,3])
}

##### proportion
# 95% confidence intervals for the correlation

set.seed(1)
data_frame_dep_v_nondep_t_x_prop_dep_mask_overlap <- subset(df_fascicle_dep_mask_overlap_t_value, select = c("depression_mask_overlap","dep_vs_healthy_t"))
results_prop <- boot(data=data_frame_dep_v_nondep_t_x_prop_dep_mask_overlap, statistic=tval,
   R=10000, formula=dep_vs_healthy_t~depression_mask_overlap)
print(results_prop) #actual 2.62, CI 0.701, 4.553)
plot(results_prop)
boot_results_prop.ci <- boot.ci(results_prop, conf=0.95, type="perc", index=1)
print(boot_results_prop.ci)
pval_prop<-boot.pval(boot_res=results_prop, theta_null = 0, type = "perc", pval_precision = NULL)
print(pval_prop) #p=0.0087

##### effect size
# 95% confidence intervals for the correlation

set.seed(1)
data_frame_dep_v_nondep_d_x_prop_dep_mask_overlap <- subset(df_fascicle_dep_mask_overlap_d_value, select = c("depression_mask_overlap","dep_vs_healthy_d"))
results_prop <- boot(data=data_frame_dep_v_nondep_d_x_prop_dep_mask_overlap, statistic=tval,
   R=10000, formula=dep_vs_healthy_d~depression_mask_overlap)
print(results_prop) #actual 2.62, CI 0.589, 4.669)
plot(results_prop)
boot_results_prop.ci <- boot.ci(results_prop, conf=0.95, type="perc", index=1)
print(boot_results_prop.ci)
pval_prop<-boot.pval(boot_res=results_prop, theta_null = 0, type = "perc", pval_precision = NULL)
print(pval_prop) #p=0.0083

##### volume
# 95% confidence intervals for the correlation
set.seed(1)
data_frame_dep_v_nondep_t_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_t_value, select = c("dep_vs_healthy_t", "volume_in_dep_mask"))

results_vol <- boot(data=data_frame_dep_v_nondep_t_x_vol_dep_mask, statistic=tval,
   R=10000, formula=dep_vs_healthy_t~volume_in_dep_mask)
print(results_vol)
plot(results_vol)
boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
print(boot_results_vol.ci)
pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
print(pval_vol) #p=1*10^-16

##### volume
#effect sizes
# 95% confidence intervals for the correlation
set.seed(1)
data_frame_dep_v_nondep_d_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_overlap_d_value, select = c("dep_vs_healthy_d", "volume_in_dep_mask"))

results_vol <- boot(data=data_frame_dep_v_nondep_d_x_vol_dep_mask, statistic=tval,
   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
print(results_vol)
plot(results_vol)
boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
print(boot_results_vol.ci)
pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
print(pval_vol) #p=1*10^-16

### with no corpus callosumm p = 1* 10^-16, CI(1.727, 5.363)
# 95% confidence intervals for the correlation

data_frame_dep_v_nondep_d_x_vol_dep_mask_no_cc <- df_fascicle_dep_mask_volume_overlap_d_value %>%
  filter(fascicle != "CC") %>%
  dplyr::select("dep_vs_healthy_d", "volume_in_dep_mask")

set.seed(1)
results_vol <- boot(data=data_frame_dep_v_nondep_d_x_vol_dep_mask_no_cc, statistic=tval,
   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
print(results_vol)
plot(results_vol)
boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
print(boot_results_vol.ci)
pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
print(pval_vol) #p=1*10^-16

```

##### PROMIS DATA

```{r compare_overall_promis_scores_by_diagnosis}

#separate out only people in depression group, and take first instance
dep_and_healthy_groups_for_ICD_analysis <-  df_people_with_promis_data_scores_added %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() #n=82

dep_vs_healthy_t <- lapply(promis_score_names, function(x) 
{
  t_test_command <- paste0("t.test(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], var.equal=F)")
  eval(parse(text=t_test_command))
})
names(dep_vs_healthy_t) <- promis_score_names
print(dep_vs_healthy_t)

#pull t-statistic
dep_vs_healthy_t_values <- sapply(dep_vs_healthy_t, function(x) x$statistic)

#convert to data frame
df_t_values <- as.data.frame(dep_vs_healthy_t_values) 
row.names(df_t_values) <- promis_score_names
print(df_t_values)

#pull p-values
dep_vs_healthy_p_values <- sapply(dep_vs_healthy_t, function(x) x$p.value)

#convert to data frame
df_dep_vs_healthy_p_values <- as.data.frame(dep_vs_healthy_p_values)
row.names(df_dep_vs_healthy_p_values) <- promis_score_names
print(df_dep_vs_healthy_p_values)
#FDR correct p-values
dep_vs_healthy_pfdr <- p.adjust(dep_vs_healthy_p_values,method="fdr")

plots <- lapply(promis_score_names, function(x) 
{

  Diagnosis <- c("MS w/Depression", "MS No Depression")
  dep_group <- eval(parse(text=paste0("dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]")))
  nondep_group <- eval(parse(text=paste0("dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]")))
  value <- c(mean(dep_group, na.rm=TRUE),
           mean(nondep_group, na.rm=TRUE))
  sem=c((sd(dep_group, na.rm = TRUE)/sqrt(sum(dep_group, na.rm = TRUE))), (sd(nondep_group, na.rm = TRUE)/sqrt(sum(nondep_group, na.rm = TRUE))))

  data <- data.frame(Diagnosis,value,sem)

  # Grouped
  barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
    geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
      geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
    scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
    ggtitle(paste0(x, " by Diagnosis")) + #, p=", round(dep_vs_healthy_t_test$p.value,3))) +
    ylab(paste0("Promis ", x)) +
    theme(title = element_text(size=14),
          axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
          axis.title.y = element_text(size=20),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
  #  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
  print(barplot) #

})

```

#### PROMIS Proportion analyses

```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis}

#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
 # filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")

promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_proportion_lost_categories

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis_dep_and_nondep}

#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")

promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_proportion_lost_categories

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis_nondep}

#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
  filter(depGroupVar == 1)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")

promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_proportion_lost_categories

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r depression_vs_healthy_overall_proportion_of_fascicles_affected_promis_dep}

#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles
#I'll read in the volume of each fascicle, and multiply it by the proportion of volume loss in each subject, thereby returing the volume of the fascicle affected

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
  filter(depGroupVar == 2)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")

promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_proportion_lost_categories

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r dimensional_promis_effects_unique_irrespective_of_dx}

#total n = 116, nothing significant

dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() #n=116

#ones to try
#PostOp.PROMIS.Mental.Score
#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"

#lm
promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_names_no_cranial_nerves

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r dimensional_promis_effects_unique_depressed_and_nondepressed}

#total n = 82, nothing significant

dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() #n=82

#ones to try
#PostOp.PROMIS.Mental.Score
#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"

#lm
promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_names_no_cranial_nerves

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r dimensional_promis_effects_unique_nondepressed}

#total n = 36, nothing significant

dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
  filter(depGroupVar == 1)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() #n=36

#ones to try
#PostOp.PROMIS.Mental.Score
#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"

#lm
promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_names_no_cranial_nerves

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r dimensional_promis_effects_unique_depressed}

#total n = 46, nothing significant

dep_and_healthy_groups_for_ICD_analysis <- df_people_with_promis_data_scores_added %>% 
  filter(depGroupVar == 2)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() #n=46

#ones to try
#PostOp.PROMIS.Mental.Score
#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"

#lm
promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_names_no_cranial_nerves

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

#### PROMIS VOLUME ANALYSES

```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis}

#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis  %>% 
 # filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")

promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_proportion_lost_categories

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})
```

```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis_dep_and_nondep}

#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis  %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")

promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_proportion_lost_categories

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis_nondep}

#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis  %>% 
  filter(depGroupVar == 1)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")

promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_proportion_lost_categories

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})
```

```{r depression_vs_healthy_overall_volume_of_fascicles_affected_promis_dep}

#goal of this segment is to look at overall lesion burden within/outside depression network, and correlate with promis scores between depression and controls, outside of fascicles

#isolate out only the depressed group and healthy group, make a column that sums up the total volume loss, only pay attention to places without cranial nerves
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
  filter(depGroupVar == 2)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_indepnet_top_quartile = 
           rowMeans(across(dep_network_names_top_quartile_no_cranial_nerves))) %>%
  mutate(total_fascicle_prop_lost_nondep_bottom_3_quartiles = 
           rowMeans(across(dep_network_names_bottom_3_quartiles_no_cranial_nerves))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

fascicle_proportion_lost_categories <- c("total_fascicle_prop_lost", "total_fascicle_prop_lost_indepnet_top_quartile", "total_fascicle_prop_lost_nondep_bottom_3_quartiles")

promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_proportion_lost_categories, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_proportion_lost_categories

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r dimensional_promis_effects_unique_irrespective_of_dx_vol}

#total n = 116, nothing significant

dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() #n=116

#ones to try
#PostOp.PROMIS.Mental.Score
#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"

#lm
promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_names_no_cranial_nerves

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r dimensional_promis_effects_unique_depressed_and_nondepressed_vol}

#total n = 82, nothing significant

dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() #n=82

#ones to try
#PostOp.PROMIS.Mental.Score
#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"

#lm
promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_names_no_cranial_nerves

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r dimensional_promis_effects_unique_nondepressed_vol}

#total n = 36, nothing significant

dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
  filter(depGroupVar == 1)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() #n=36

#ones to try
#PostOp.PROMIS.Mental.Score
#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"

#lm
promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_names_no_cranial_nerves

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```

```{r dimensional_promis_effects_unique_depressed_vol}

#total n = 46, nothing significant

dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicle_volumes_full_w_promis %>% 
  filter(depGroupVar == 2)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() #n=46

#ones to try
#PostOp.PROMIS.Mental.Score
#[1] "Quality_of_Life"                  "Physical_Health"                  "Mental_Health_and_Mood"           "Social_Activities_Satisfaction"  
# [5] "Carrying_Out_Social_Activities"   "Carrying_Out_Physical_Activities" "Emotional_Problems"               "Fatique_Average"                 
# [9] "PostOp.PROMIS.Physical.Score"     "PostOp.PROMIS.Mental.Score"

#lm
promis_lm <- lapply(promis_score_names, function(promis_score) {

  fascicle_lm <- lapply(fascicle_names_no_cranial_nerves, function(fascicle) 
  {
    lm(substitute(fascicle ~ promis_score, list(fascicle = as.name(fascicle), promis_score=as.name(promis_score))), 
       data = dep_and_healthy_groups_for_ICD_analysis)
  })
  names(fascicle_lm) <- fascicle_names_no_cranial_nerves

  #anova
  fascicle_anova <- lapply(fascicle_lm, anova)

  #fdr corrected
  fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
  print(fascicle_anova_fdr)
})

```