### 2022-08-26 msdepression/scripts/promis_score_cleaning_2010-2022.Rmd
Add files via upload

---
title: "Cleaning PROMIS scores"
author: "Erica Baller"
date: '2022-08-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Goal of this script is to clean the promis scores and assemble them in a better table format for later processing.

Number of non NA values for each of the promis scores:

General_Health                  Quality_of_Life                  Physical_Health 
                            1694                             1693                             1503                             1503 
          Mental_Health_and_Mood   Social_Activities_Satisfaction   Carrying_Out_Social_Activities Carrying_Out_Physical_Activities 
                            1503                             1503                             1501                             1499 
              Emotional_Problems                  Fatique_Average     PostOp.PROMIS.Physical.Score       PostOp.PROMIS.Mental.Score 
                            1499                             1499                             1440                             1442

```{r score_cleaning}
#before doing this, must save the promis score spreadsheet, /Users/eballer/BBL/msdepression/data/dac/copyofpromisflowsheetdata2010-2022.xlsx in two separate csv files (because it has two tabs). This was done manually and separated by date. New files are:
#/Users/eballer/BBL/msdepression/data/dac/promis_data_2015.csv
#/Users/eballer/BBL/msdepression/data/dac/promis_data_2016-2022.csv

#read in spreadsheets
promis_data_2015 <- read.csv("/Users/eballer/BBL/msdepression/data/dac/promisflowsheetdataupdated_good_physical_health_scores_20220824_sheet1.csv", sep = ",", header = T) #n=2041
promis_data_2016_2022 <- read.csv("/Users/eballer/BBL/msdepression/data/dac/promisflowsheetdataupdated_good_physical_health_scores_20220824_sheet2.csv", sep = ",", header = T)#n=

promis_data_combined <- rbind(promis_data_2015, promis_data_2016_2022) #n=6916

#promis score names, from 7th column all the way to the end of the spreadsheet, starting with general health to PostOp.PROMIS.Mental.Score

promis_score_names <- names(promis_data_combined[7:dim(promis_data_combined)[2]-1])

##### cleaning ####

#total scores n=1694
promis_data_cleaned <- promis_data_combined %>%
  mutate(across(everything(), gsub, pattern = "NULL", replacement = "NA")) %>% #replace null with NA for easier use
  filter(RECORDED_TIME != "NA") %>% #remove nulls in Recorded time
  mutate(date_promis_collected = as.Date(RECORDED_TIME, format = "%m/%d/%y")) %>% 
  mutate(date_promis_collected = gsub(date_promis_collected, pattern = "-", replacement = "")) %>% #get time into better format, now matches other spreadsheet
  mutate(year_promis_collected = gsub(date_promis_collected, pattern = "\\d{4}$", replacement = "")) %>% #make new column that just as date and change to numeric
  mutate(across(everything(), gsub, pattern = "\\s*-.*", replacement = "")) %>% #\\s = whitespace, .* any character, replace with blank, to clean the text up
  mutate_at(promis_score_names, as.numeric) %>% #convert columns from character to numeric
  mutate(year_promis_collected = as.numeric(year_promis_collected)) %>% #convert to numeric
  mutate(RECORDED_TIME = as.Date(RECORDED_TIME, format = "%m/%d/%y")) #turn back into date format

#histos
hist(promis_data_cleaned$year_promis_collected) #want to see spread of when scores collected

for (i in promis_score_names) {
  eval_for_parse <- paste0("hist(promis_data_cleaned$", i, ")")
  eval(parse(text=eval_for_parse))
}

#look at number of samples in each column promis_data_cleaned
colSums(!is.na(promis_data_cleaned))

unique_individuals <- promis_data_cleaned %>% 
  group_by(EMPI) %>% 
  slice(1) %>%
  ungroup #n = 501

total_num_promis_scores <- dim(promis_data_cleaned)[1] #n= 1694

num_unique_indiv <- dim(unique_individuals)[1] #n=501

saveRDS(promis_data_cleaned, file = paste0("/Users/eballer/BBL/msdepression/data/dac/promis_data_20220824_fixed_physical_health_cleaned_n", total_num_promis_scores, "_unique_indiv_n", num_unique_indiv, "_2015-2022.rds"))
```