### 2022-09-02 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_just_volumes_started_20220824.Rmd
massive cleanup - just need to fix sensis

---
title: "MS Depression Paper, started 2022 08 24"
author: "Erica Baller"
date: '2022-08-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(boot)
library(boot.pval)
library(lsr)
library(sjPlot)
library(coin)
library(rstatix)

#functions
source("~/BBL/msdepression/scripts/msdepression_functions.R")
#source("~/BBL/msdepression/scripts/wilcox_effsize.R")
```

Read in Data and do some preprocessing

```{r read_in_df_and_subset}
homedir <- "/Users/eballer/BBL/msdepression/"

data <- read.csv(paste0(homedir, "/data/drugs_data/parsable_msdepression.csv"), sep = ",", header = TRUE) #n=16830, only has people with MS ICD10 code G35 and seen by MS provider

columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE","hemoglobin", "WBC","RBC","B12","FOLATE", "TSH", "RPR","CSFAPPEAR1","CSFAPPEAR4","CSFCOLOR1",       "CSFCOLOR4","CSFTUBE","CSFTUBE4","VitD","PHQ.2","PHQ.9")

##########################
### some preprocessing ###
##########################
#we need to keep scans from people seen by an MS provider (n = 17067) who have an MS diagnostic code (n=16,830)
#we make a bunch of columns from character to integer, binarize sex and race, and put date into a nice format YYYYMMDD, the %m%d%y indicates what it started out as

#goal is to keep people who were seen by 
 data_empi_acc_f_phq <- data %>% 
   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = "")) %>%
   mutate(BEGIN_EXAM_DTTM = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% #make sure to actually save it as date object for later processing
   mutate(EMPI = as.factor(EMPI)) %>%
   mutate(On.Psych.Meds = ifelse(On.Psych.Meds == "True", 1, 0)) %>%
   mutate(On.Antidepressants = ifelse(On.Antidepressants == "True", 1, 0)) %>%
   mutate(Has.PHQ2 = ifelse(Has.PHQ2 == "True", 1, 0)) %>%
   mutate(Has.PHQ9 = ifelse(Has.PHQ9 == "True", 1, 0)) %>%
   mutate(Has.depdx = ifelse(grepl("F3", ICD10), 1, 0)) %>%
   mutate(PHQ.2_modsev_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 >= 3), 1, 0)) %>%
   mutate(PHQ.9_modsev_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 >= 10), 1, 0)) %>%
   mutate(PHQ.2_mild_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 > 0 & PHQ.2 < 3), 1, 0)) %>%
   mutate(PHQ.9_mild_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 > 0 & PHQ.9 < 10), 1, 0)) %>%
   mutate(PHQ.2_zero = ifelse((!is.na(PHQ.2) & PHQ.2 == 0), 1, 0)) %>%
   mutate(PHQ.9_zero = ifelse((!is.na(PHQ.9) & PHQ.9 == 0), 1, 0)) %>%
   mutate(dep_by_dx_phq = ifelse((Has.depdx | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(dep_by_dx_phq_antidep = ifelse((Has.depdx | On.Antidepressants | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(true_healthy = ifelse((!Has.depdx & !On.Psych.Meds & (PHQ.2_zero | PHQ.9_zero)), 1, 0)) %>%
   mutate(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds = ifelse(dep_by_dx_phq_antidep | true_healthy, 1, 0)) %>%
   rowwise(ACCESSION_NUM) %>%
   mutate(depGroupVar =
            sum(c(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds,dep_by_dx_phq_antidep))) %>%#you get an extra point if you are in the depression group, so the end result is dep = 2, healthy = 1, 0 for exclude
   ungroup()

##########################################################
#             Read in Fascicle info             #
##########################################################

fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77, drop cranial nerves

#add suffix to each measure
fascicle_names_w_prop_suffix <- gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE, x = fascicle_names_no_cranial_nerves)
fascicle_names_w_vol_suffix <- gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE, x = fascicle_names_no_cranial_nerves)

########
#fascicle volumes in depression network
########

fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")

##########################################################
#             Read in Lesion Volumes                     #
##########################################################

# contains the volume of lesioned voxels, i.e. a sum of all the areas of each individual's binary mask that are 1

lesion_volumes <- read.csv(paste0(homedir, "results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes"), sep = ",", header = TRUE)

##########################################################
#             Merge with data_empi              #
##########################################################
#merge whole data set with fascicle proportions
df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962

#merge with the df that has lesion volumes, 1 per perosn
df_demo_and_fascicles <- merge(df_demo_and_fascicles, lesion_volumes, by = c("EMPI", "EXAM_DATE")) #n = 2962

##########################################################
#  Add columns for volume and change proportion names    #
##########################################################

#read in fiber volume values
fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)

#remove cranial nerves
fiber_volume_values_no_cranial_nerves <- fiber_volume_values %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

#extract data frame that only includes the fascicle proportion lost
df_fascicles_prop <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)]
df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)]

#multiply each row by the vector of total volumes to extract volume of disease within fascicle
df_fascicles_volumes <- sweep(x = df_fascicles_prop, MARGIN = 2,                        # Sweep with Complex STATS
                  STATS = fiber_volume_values_no_cranial_nerves$volume_full, FUN = "*") %>%
  mutate_if(is.numeric, round) #round all values

# change name of df_fascicle_volumes to have the suffix, and add a suffix for the proportion measures
names(df_fascicles_prop) <- fascicle_names_w_prop_suffix
names(df_fascicles_volumes) <- fascicle_names_w_vol_suffix

#combine back with data_fram
#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
df_demo_and_fascicles_prop_and_vol <- cbind(df_everything_else, df_fascicles_prop, df_fascicles_volumes )

```
PROMIS DATA HANDLING to link a person to their most proximal PROMIS

```{r promis_data_handling}
##########################################################
#            Read in PROMIS Data              #
##########################################################
promis_data <- readRDS(paste0(homedir, "/data/dac/promis_data_20220824_fixed_physical_health_cleaned_n1813_unique_indiv_n532_2015-2022.rds"))
promis_score_names <- names(promis_data[7:16])

df_people_with_promis_data <- df_demo_and_fascicles_prop_and_vol %>%
  filter(EMPI %in% promis_data$EMPI) #n=398

unique_empis_in_promis_group <- unique(df_people_with_promis_data$EMPI) #n=116

#go through each person w/promis data and append the columns to their row that reflects the closest collected PROMIS

#function, take an empi and date, returns the row of data w/promis scores
get_closest_promis_to_ms_exam_date <- function(subj_empi, subj_exam_date) {
  df_closest_promis <- promis_data %>%
    filter(EMPI == subj_empi) %>%
    mutate(distance_from_scan = abs(RECORDED_TIME-subj_exam_date)) %>%
    filter(abs(RECORDED_TIME-subj_exam_date)==min(abs(RECORDED_TIME-subj_exam_date))) %>%
    slice(1)
  return(df_closest_promis)
}

#make dataframe to store promis info
promis_info_closest_proximity <- data.frame(matrix(NA, nrow=dim(df_people_with_promis_data)[1], ncol=(dim(promis_data)[2] + 1)))
names(promis_info_closest_proximity) <- c(names(promis_data), "distance_from_scan")

for (line in 1:dim(df_people_with_promis_data)[1]) {
  #pull out subject empi and exam date
  subj_empi <- df_people_with_promis_data$EMPI[line]
  subj_exam_date <- df_people_with_promis_data$BEGIN_EXAM_DTTM[line]

  #returns whole line, promis scores, id, empi, everything
  closest_promis_data <- get_closest_promis_to_ms_exam_date(subj_empi = subj_empi, subj_exam_date = subj_exam_date)

  #store as new line
  promis_info_closest_proximity[line,] = closest_promis_data
}

#drop unnecessary columns c("DE_PAT_ID", "EMPI", "HUP_MRN", "SEX", "RACE")
promis_info_closest_proximity_just_scores <- promis_info_closest_proximity[,6:dim(promis_info_closest_proximity)[2]]

#combine with df_people_with_promis_data
df_people_with_promis_data_scores_added = cbind(df_people_with_promis_data, promis_info_closest_proximity_just_scores)

#summarize the distance from scans
summary(df_people_with_promis_data_scores_added$distance_from_scan/364) #mean 2.40 ys, median 2.10 years, range 0.003 y to 6.83 years

hist(df_people_with_promis_data_scores_added$distance_from_scan/364, 
     main = "Hisogram of years between Promis and Scan",
     xlab = "Years from Scan") #convert to years

```

Demographics
```{r demographics}

#this will display a demographics table comparing age, race, sex between healthy and depressed groups, using the first presentation to treatment

df_good_mimosa_unique_dep_and_healthy <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique_dep_and_healthy)

#just good mimosa w/promis
df_people_with_promis_data_scores_added_unique <- df_people_with_promis_data_scores_added %>%
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms_w_promis(data_frame = df_people_with_promis_data_scores_added_unique)

```

Define Depression Network by Volume of Overlap
```{r visualize_depression_network_volume}
###################################################
### Defining and Visualizing Depression Network ###
###################################################

#This chunk reads in a file where columns are fascicle, num_voxels_total, non_zero_voxels_in_dep_map, prop_in_mask
# by the end, the depression mask will be defined
# We consider fascicles in the depression network to be the ones that share the most VOLUME with the depression mask
# Of course, it is possible that huge fascicles contribute a ton to the depression network, but also the non-depression network
# And to get at specificity (i.e. which fibers most exclusively send information within the depression network), we could look at proportion of the fiber (volume that intersects / total volume of the fascicle). But then you have a situation where tiny fibers get lots of credit. We have actually tried this as well and the findings aren't that different. 
#But for the sake of this paper, will do volume because the question is all volume based (volume of lesions to volume of fascicles)

######## Histograms for depression network
#first set the values to be the volumes
fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_dep %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

#set NAs to zero
fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map[is.na(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map)] = 0

#sort based on overlap with depression depression map
fascicle_volumes_dep_network_no_cranial_nerves$fascicle <- factor(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, levels = fascicle_volumes_dep_network_no_cranial_nerves$fascicle[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map)])

#figure out how many fascicles are in top quartile
quartile <- round((dim(fascicle_volumes_dep_network_no_cranial_nerves)[1] / 4),0)

#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take names of the top quartile
dep_network_names_top_quartile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile]

#get names of bottom 3 quartiles

#get a network that includes all of the fascicles NOT in dep 25% network
dep_network_names_bottom_3_quartiles_no_cranial_nerves <- fascicle_volumes_dep_network_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves)]

#### Get sums of the volumes of the fascicles in each network, and total sum

#top quartile, depression network
dep_network_total_volume_top_quartile_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves)])

#bottom 3 quartiles, nondepression network
bottom_75_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves)])

#### Calculate total volume in all non-cranial_nerve_network
total_volume_all_full_fascicles <- dep_network_total_volume_top_quartile_no_cranial_nerves + bottom_75_total_volume_no_cranial_nerves

###

plot_all_fascicle_volumes_dep_network_no_cranial_nerves <-ggplot(data = fascicle_volumes_dep_network_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map))+
  ggtitle("Volume Overlap per Fascicle 3_09") + 
  geom_bar(stat="identity", fill="006300") +
  xlab("Fascicle") +
  ylab("Overlap with Depression Mask") +
  ylim(0,40000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicle_volumes_dep_network_no_cranial_nerves)

### Change names to add suffix
#replace each fascicle with vol suffix
dep_network_names_top_quartile_no_cranial_nerves <- dep_network_names_top_quartile_no_cranial_nerves %>% 
  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix

dep_network_names_bottom_3_quartiles_no_cranial_nerves <- dep_network_names_bottom_3_quartiles_no_cranial_nerves %>%
   gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix
```

Compare overall volume of the lesions, irrespective of fascicle involvement, between diagnostic groups

```{r compare_overall_volume_of_lesions_dep_vs_non_dep}

#separate out only people in depression group, and take first instance
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# Assessing Normality w/histograms and Shapiro_Wilk test for normality
hist(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])
non_dep_lesion_vol_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]) #W=0.77, p = 8.2 * 10^-14, NON-NORMAL

hist(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
dep_lesion_vol_shapiro <- hist(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
shapiro.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]) #W=0.80, p = 2.2 * 10^-16, NON-NORMAL

dep_vs_healthy_wilcoxon_lesion_volume_results <-  wilcox.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired = F) #p=0.036; depressed = 15,871; nondepressed = 12,886

dep_vs_healthy_wilcoxon_lesion_volume_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = volume_of_mimosa_lesions ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)

#extract overall volume and diagnosis
df_lesion_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
  dplyr::select(volume_of_mimosa_lesions,depGroupVar) %>%
  filter(depGroupVar != 0) %>% 
  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))

violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, fill=depGroupVar)) +#, fill=depGroupVar)) + 
  geom_violin() + 
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
  ylab("Volume of Lesions") +
  theme(plot.title = element_text(size=16, hjust = 0.8),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
print(violin_lesion_volume) #p=0.067

```

```{r depression_vs_healthy_overall_volume_of_fascicles_affected}
## compare disease within and outside depression network, between diagnoses, and interactions

###############################
### Constructing Data Frame ###
###############################

#make a data frame for only unique individuals, and get means of overall volume of disease across all fascicles, and within/outside dep net
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>%

  #sum of total volume lost and proportion of volume lost 
  mutate(sum_fascicle_vol_lost = rowSums(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  mutate(proportion_volume_lost_per_total_network_size_sum = sum_fascicle_vol_lost/total_volume_all_full_fascicles) %>%

  #sum of total volume lost in depression network and proportion of volume lost in depression network (volume lost/network volume size) 
  mutate(sum_fascicle_vol_lost_indepnet_top_quartile = 
           rowSums(across(all_of(dep_network_names_top_quartile_no_cranial_nerves)))) %>%
  mutate(proportion_volume_lost_dep_net_by_network_size_sum = sum_fascicle_vol_lost_indepnet_top_quartile /dep_network_total_volume_top_quartile_no_cranial_nerves) %>%

  #sum of total volume lost in nondepression network and proportion of volume lost in depression network (volume lost/network volume size) 
  mutate(sum_fascicle_vol_lost_nondep_bottom_3_quartiles = 
           rowSums(across(all_of(dep_network_names_bottom_3_quartiles_no_cranial_nerves)))) %>%
  mutate(proportion_volume_lost_nondep_net_by_network_size_sum = sum_fascicle_vol_lost_nondep_bottom_3_quartiles /bottom_75_total_volume_no_cranial_nerves) %>%

  ungroup()

###################

##### Assessing Normality w/histograms and Shapiro_Wilk test for normality ####

### Test for normality within each network, within each diagnosis

disease_overall_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum) #W=0.97, p = 3.4 * 10^-7, NON-NORMAL

disease_dep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum) #W=0.97, p = 5.0 * 10^-7, NON-NORMAL

disease_nondep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum) #W=0.97, p = 1.9 * 10^-7, NON-NORMAL

disease_dep_dx_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]) #W=0.97, p = 0.0002 NON-NORMAL

disease_nondep_dx_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]) #W=0.96, p = 0.0001 NON-NORMAL

## All are non-normal, will use wilcoxon for comparisons

### Within/outside depression network, p = 2.2*10^-16, dep net prop = 0.400, non dep = 0.339
msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum,paired = TRUE)

### Diagnostic effects

#proportion volume lost over whole network = W=14995, p=0.04, dep = 0.39, nondep = 0.30
proportion_volume_lost_over_whole_brain_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)

#effsize 0.11
proportion_volume_lost_over_whole_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_per_total_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)

#proportion volume lost within depression network, W = 14941, p-value = 0.03; dep = 0.42, nondep = 0.38
proportion_volume_lost_dep_net_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)

#effsize 0.11
proportion_volume_lost_over_dep_net_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_dep_net_by_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)

#proportion volume lost outside depression network W = 15120, p-value 0.05; dep = 0.36, nondep = 0.33
proportion_volume_lost_nondep_net_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)

#effsize 0.10
proportion_volume_lost_over_nondep_net_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_nondep_net_by_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)

### Interaction Network x Diagnosis
scale=1

df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

#eff size = 0.73
msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions, formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)

#diagnosis: Dep > nondep T = 2.285, p = 0.023; Network dep>nondep = T=16.6, p = 2*10^-16; intx = T = 2.21, p = 0.027
diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
summary(diagnosis_by_network_interaction_lmer)

```

``` {r plots_for_dx_net_intx}
########## PLOTS #############

#Within vs outside network
dep_net_group <- c(rep("Dep_net", times = dim(dep_and_healthy_groups_for_ICD_analysis)[1]), rep("Non_dep_net",times =dim(dep_and_healthy_groups_for_ICD_analysis)[1]))
net_values <- c(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum)

newdf <- data.frame(dep_net_group, net_values)

violin_within_vs_outside_network_sum_proportion_scaled_by_network_size <- ggplot(newdf, aes(x=dep_net_group, y=net_values, fill=dep_net_group)) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "Dep Net", "2" = "Non-Dep Net")) + 
  ggtitle(paste0("Proportion Fascicle Volume Affected by Network")) +
  ylab("Proportion (sum total volume lost/Net Vol) ") +
  ylim(0,1) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(violin_within_vs_outside_network_sum_proportion_scaled_by_network_size) #

#diagnosis

df_proportion_network_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
dplyr::select(proportion_volume_lost_per_total_network_size_sum, depGroupVar) %>%
  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))

violin_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis, aes(x=depGroupVar, y=proportion_volume_lost_per_total_network_size_sum, fill=depGroupVar)) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "NonDepressed", "2" = "Depressed")) + 
  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
  ylab("Proportion (sum total volume lost/Net Vol) ") +
  ylim(0,1) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(violin_dep_vs_nondep_scaled_by_network_size) #

#interaction
diagnosis_by_network_interaction_lmer_plot <- plot_model(diagnosis_by_network_interaction_lmer, 
                                                         type = "int",
                                                         axis.lim = c(0.2,0.5),
                                                         colors = c("#ff00ff", "#3df2f2"),
                                                         line.size = 1.3,
                                                         dodge=0.2,
                                                         axis.title = c("Diagnosis", "Proportion Affected"),
                                                         title = "Diagnosis x Depression Network Disease")+

  theme(axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        aspect.ratio = 3/2,
        panel.background = element_rect(colour = "white", fill = "white"))

print(diagnosis_by_network_interaction_lmer_plot)

```

``` {r effect_size_regression}
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# lapply to get results of t-test in each fascicle
#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_w_vol_suffix, function(x) 
{
  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
  eval(parse(text=text_for_eval))
})
names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_w_vol_suffix

#convert to data frame
df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)

```

``` {r bootstrap}
#### bootstrap

######### functions
tval <- function(formula, data, indices) {
  df <- data[indices,] # allows boot to select sample
  fit <- lm(formula, data=df)
  return(summary(fit)$coefficients[2,3])
}

#effect sizes
# 95% confidence intervals for the correlation
set.seed(1)
data_frame_dep_v_nondep_d_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_overlap_d_value, select = c("dep_vs_healthy_d", "volume_in_dep_mask"))

results_vol <- boot(data=data_frame_dep_v_nondep_d_x_vol_dep_mask, statistic=tval,
   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
print(results_vol)
plot(results_vol)
boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
print(boot_results_vol.ci)
pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
print(pval_vol) #p=1*10^-16

```

``` {r plot_scatter}
#######################
### plot differences ##
#######################

#combine data frames, column 1 has names of fascicles, column 2 has Volume of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle

df_fascicle_dep_mask_volume_overlap_d_value <- data.frame(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, df_d_values)
names(df_fascicle_dep_mask_volume_overlap_d_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_d")

# plot fascicle overlap x Eff size value
#statistic, p=001
dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value )

scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect Size (D)") + 
  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", pval_vol)) + 
                 #round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
#  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network")) + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text = element_text(size=12, color = "black"),
        axis.title = element_text(size=14, color = "black"),
        axis.line = element_line(color="black", size=1.3),
        panel.background = element_blank())
print(scatter_d_val_by_dep_net_overlap)

df_for_making_colors <- df_fascicle_dep_mask_volume_overlap_d_value %>% dplyr::select(fascicle, dep_vs_healthy_d)
write.table(x = df_for_making_colors, file = paste0(homedir, "results/fascicle_and_effect_size_for_volxdx_analysis_n77.csv"), row.names = F, col.names = F, quote = F, sep = ",")

```

```{r sensitivity analyses}
tertile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 3),0)
quintile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 5),0)

dep_network_names_top_tertile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:tertile] #n=26

dep_network_names_top_quintile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quintile] #n=15

dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves_by_vol)]

dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves_by_vol)]

top_tertile_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves_by_vol)])

top_quintile_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves_by_vol)])

bottom_tertiles_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol)])

bottom_quintiles_total_volume_no_cranial_nerves<-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol)])

dep_and_healthy_groups_for_ICD_analysis <- dep_and_healthy_groups_for_ICD_analysis %>%

  #tertile and quintiles
  mutate(sum_fascicle_vol_lost_top_tertile = 
           rowSums(across(dep_network_names_top_tertile_no_cranial_nerves_by_vol))) %>%
  mutate(proportion_volume_lost_top_tertile_by_network_size_sum = sum_fascicle_vol_lost_top_tertile /top_tertile_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_bottom_tertiles = 
           rowSums(across(dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol))) %>%
  mutate(proportion_volume_lost_bottom_tertiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_tertiles /bottom_tertiles_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_top_quintile = 
           rowSums(across(dep_network_names_top_quintile_no_cranial_nerves_by_vol))) %>%
  mutate(proportion_volume_lost_top_quintile_by_network_size_sum = sum_fascicle_vol_lost_top_quintile /top_quintile_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_bottom_quintiles = 
           rowSums(across(dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol))) %>%
  mutate(proportion_volume_lost_bottom_quintiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_quintiles /bottom_quintiles_total_volume_no_cranial_nerves) %>%

#### Save a data frame that contains the mean volume of disease within network, scaled by the size of the network, for visualization

diagnosis_vector <- c(1,2,1,2) #1 is healthy, 2 is depressed
network_vector <- c(1,1,2,2) #1 is depression network, 2 is non_depression network
mean_diagnosis_by_network_vector <- c(mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]))

#df_means_volume_scaled_by_network_size <-data.frame(diagnosis_vector, network_vector, mean_diagnosis_by_network_vector)
#names(df_means_volume_scaled_by_network_size) <- c("diagnosis", "network", "mean")
#write.table(df_means_volume_scaled_by_network_size, paste0(homedir, "/results/df_means_of_volumes_scaled_by_network_size_dx_net_for_color.csv"), row.names=F, col.names = T, quote = F, sep=",")

#### tertile/quintile

#p=2.2*10^16, top tertile prop vol affected = 0.420, bottom 2 tertiles = 0.298
msdisease_within_versus_outside_depression_network_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum,paired = TRUE)

#p=0.0328, dep = 0.4407, nondep = 0.387
msdisease_diagnostic_effects_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])

#p=0.058, dep = 0.31, nondep= 0.28
msdisease_diagnostic_effects_wilcoxon_bottom_tertiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])

#p=2.2*10^-16, top quintile prop vol affected = 0.419, bottom 4 quintiles = 0.322
msdisease_within_versus_outside_depression_network_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum,paired = TRUE)

#p=0.0325, dep = 0.4411, nondep = 0.386
msdisease_diagnostic_effects_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])

#p=0.048, dep = 0.34, nondep = 0.30
msdisease_diagnostic_effects_wilcoxon_bottom_quintiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])

```