### 2022-09-16 msdepression/scripts/visualization/vol2surf_mapping.Rmd
Volume to surface mapping from Val

---
title: "Project an MNI volume to the fslr surface for visualization"
author: "Valerie Jill Sydnor, adapted by Erica Baller"
date: "9/16/2022"
output: html_document
---

```{r setup, include=FALSE}
library(ciftiTools)
ciftiTools.setOption('wb_path', '/Users/eballer/Software/workbench/')
#set the path to your installation of connectome workbench as 'wb_path'
library(rgl) #to use ciftiTools graphics
library(rmarkdown)
library(gifti) #to read in your surface giftis 
```

Map volume data to left and right cortical surfaces
```{r}
#left hemisphere
command1=sprintf("-volume-to-surface-mapping /Users/eballer/BBL/msdepression/templates/harvard_depression/surfaces/DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_3_09.nii /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii /Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_l.shape.gii -trilinear")
ciftiTools::run_wb_cmd(command1, intern = FALSE, ignore.stdout = NULL, ignore.stderr = NULL)

#right hemisphere
command2=sprintf("-volume-to-surface-mapping /Users/eballer/BBL/msdepression/templates/harvard_depression/surfaces/DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_3_09.nii /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii /Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_r.shape.gii -trilinear")
ciftiTools::run_wb_cmd(command2, intern = FALSE, ignore.stdout = NULL, ignore.stderr = NULL)
```

Read in your surface giftis
```{r}
test_l <- read_gifti("/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_l.shape.gii")
test_r <- read_gifti("/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_r.shape.gii")
```

Combine your giftis into one cifti surface (for visualization with ciftiTools)
```{r}
test.ciftimap <- as_cifti(cortexL = test_l$data$normal, cortexR = test_r$data$normal) #the datas are stored in file$data$normal
test.ciftimap$data$cortex_left[test.ciftimap$data$cortex_left > 0] <- 1
test.ciftimap$data$cortex_right[test.ciftimap$data$cortex_right > 0] <- 1

```

You can also save out this cifti for viewing in connectome workbench
```{r}
write_cifti(test.ciftimap, "/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf.dscalar.nii")
```

Choose what surface you want to visualize your data on (midthickness, inflated, sphere, etc.)
```{r}
#lets use midthickness
surfL_name <- read_surf("/Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii")
surfR_name <- read_surf("/Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii")

#gotta add these surfs to your cifti from the step above
test.ciftimap <- add_surf(test.ciftimap, surfL=surfL_name, surfR=surfR_name)
```

Visualize!
```{r, rgl=TRUE, format="png"} 
view_cifti(test.ciftimap, colors=c("magma"), widget=TRUE)
```