### 2022-10-06 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_2022092022_post_replication.Rmd
Add files via upload

---
title: "MS Depression Paper, post replication, started 2022 09 21"
author: "Erica Baller"
date: '2022-08-24'
output:
  html_document: default
  pdf_document: default
---

## MS Depression

Welcome to the MS Depression Data! This project was designed to ask a relatively simple question - are patients with MS who are depressed more likely to have lesions in fascicles that connect areas previously associated with depression. The literature are sparse in this domain. It is very clear that depression occurs in up to 50% of patients with MS, and some studies have suggested that it relates to lesion burden. But to my knowledge there are no papers that evaluate how lesions disrupt white matter tracts that serve as the backbone for functional networks. Over the course of the next few chunks, I'll walk through our analysis. It combines imaging data from research grade scans acquired at 3T as part of a clinical protocol (T1 and T2 Flair), and clinical data from the electronic medical record.

The first chunk here is reading in appropriate data frames and preprocessing. A few big data frames are read in. The first is the demographic, drug, lab info for the MS patients (from the DAC pull). The second is the data frame of fascicle proportion affected for people with good MIMoSA (# volume affected/total volume of fascicle). The most important preprocessing occurs in a series of mutate commands. These essentially add new columns to the data frame that can later be used. The thrust of these mutate commands is to define who has depression and who does not.

Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion. Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses. For these preliminary analysis, our inclusion/exclusion are more stringent. Of note, the graph below is irrespective of whether they had good/bad MIMoSA. As such, the MIMoSA subset is smaller.

<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>

### Inclusion/Exclusion

#### Inclusion MS Criteria (dataframe read in here has already been filtered down using the exclusions below previously)

  1) Scanned under MS protocol (42K)

  2) MS Diagnosis (17K)

  3) Seen by MS doctor (16K)

#### Inclusion criteria for Depressed Group (one of the three criteria must be met)

  1) Have an ICD10 F3* code (in our sample, people have F32 (depressive episode), F33 (major depressive disorder), and F34 (Persistent mood [affective] disorders)) 
        - there are NO participants with bipolar disorder, manic episode, or unspecified mood/affective disorder
        - while participants may have been diagnosed with F06.3 (Depressive disorder due to known physiological condition, with depressive features), they must ALSO have had an F3(2-4) condition

  2) Screened positive for depression on PHQ2 or PHQ9 at any point

  3) Prescribed antidepressant medications per NAMI website
        - https://www.nami.org/About-Mental-Illness/Treatments/Mental-Health-Medications

#### Inclusion criteria for Nondepressed Group (ALL three criteria must be met)

  1) No F* diagnoses

  2) At least 1 PHQ2 or PHQ9 with a documented 0

  3) Free of psychiatric medications per NAMI website

#### EMR Exclusions

  If you received the MS protocol or had a diagnosis of MS but never actually saw an MS provider, we cannot be sure that the diagnosis of MS is truly valid, as it is hard to diagnose anyway.
  If you did not have a diagnosis of depression and never completed a PHQ2 or 9, and had no history of medications, we could not safely confirm nor deny the presence of depression. These subjects were excluded from future analyses. We could consider adding these to the healthy group maybe, but not too inclined to do that right now.

#### Imaging exclusion

  We are currently only using people who had good MIMoSA (75 or 100). That severely limits our numbers, but that is okay! When we re-QC, we will probably get a lot more into this group. We are specifically using their FIRST scan.

### Group validation and depression network definition

#### Functional Impairment Scores

  To validate our grouping, we wanted to compare symptom burden between patients with depression and nondepressed comparators. One challenge of EMR research is that we are limited to what has been coded in particular fields, and could not readily assess our patients, or access their notes. As PHQ2 and 9 were used in dignostic categorization, they, by default differ between groups and are less informative. MS impairment is often assessed and recorded via EDSS, but this is not currently coded in the EMR. However, the Penn EMR has started recording Patient-Reported Outcomes Measurement Information System (PROMIS). Per the NIH, "PROMIS addressed a need in the clinical research community for a rigorously tested patient reported outcome (PRO) measurement tool that uses recent advances in information technology, psychometrics, and qualitative, cognitive, and health survey research to measure PROs such as pain, fatigue, physical functioning, emotional distress, and social role participation that have a major impact on quality-of-life across a variety of chronic diseases." PROMIS scores, for reference, are scored where higher numbers mean better functioning, and lower numbers mean worse functioning.

  A small subset of our participents (36 in non-depressed group and 49 in depressed group) had PROMIS scores. For participants who had multiple PROMIS scores, the score that was analyzed was the closest in time proximity to the day of the scan. PROMIS score cleaning and linking of score to scan is performed in a separate markdown **(promis_score_cleaning_2010-2022.Rmd). **

  Based on previous MS literature, we hypothesized that our patients with depression would overall show more functional impairment than patients without depression.

#### Depression Network Construction

  A note on the "depression network." This network was made by Shan Siddiqi in Michael Fox's group by doing a conglomeration of TMS datasets, stroke datasets, Nature Human Behavior 2021 (https://www.nature.com/articles/s41562-021-01161-1). It is an unthresholded mask (he initially sent one where the unit is r, but then sent us one where unit is T). 
  Per Siddiqi's article: The location of each lesion or brain stimulation site (Fig. 2a–c, top panels) was mapped to an underlying brain circuit using a large normative connectome database (n=1000) and previously validated methods (Fig. 2a–c, bottom panels)5. The normative connectome was used to estimate connectivity of each lesion or stimulation site to every voxel in the brain. At each voxel, a Pearson r-value was computed for the correlation between depression score and lesion or stimulation site connectivity to that voxel (Fig. 2a–c, right panels), yielding a population-derived “circuit map” for each of the 14 datasets (Supplementary Figure 1). We generated a combined depression circuit map by taking the mean of all 14 circuit maps, weighted by the sample size of each dataset (Fig. 5a). Of note, lesions are expected to positively correlate with symptoms, whereas the inverse of tms sites and was (so connectivity in these areas was inverted).

  To use it, I thresholded the clusters (3.09), binarized it and then used it as an ROI and calculated, per fascicle, the volume occupied by the fascicle that intersected with the depression mask. Most fascicles were either entirely overlapping or entirely outside the depression mask, though the range of overlap was anywhere from 1-30% (volume of fascicle overlap/total volume of fascicle). This required us to make some choices. We considered an option where the "depression mask" was anywhere that ever overlapped, and non-depression were areas where there was no overlap. We also looked at quartiles, and settled on the top 25% (top quartile), i.e. the top 25% of fascicles with the highest volume of network overlap were considered in the depression network. Everything outside of that was considered "non_depression" network, or nondep_net. Cranial nerves are too small to reliably be assess with our method, and so they were removed from the analysis.In total, 77 fascicles were evaluated.

### Analyses
  Demographic differences between groups were assessed with T-tests and chi-square for categorical variables w/the command CreateTableOne.

#### *Disease burden measures per individual (done in the shell):*

    For each individual, the number of 1s in the MIMoSA binary mask were summed to get a measure of overall lesion burden (to be distinguished from fascicle burden).

    To assess fascicle burden for each individual, streamline filtering was performed.

    Streamline filtering is an interative process performed in DSI studio. For each individual, the MIMoSA binary map was considered a region of interest. For each of the 77 fascicles, streamlines that ran through the lesion were "filtered" or kept, whereas the fascicles that avoided the MIMoSA mask were eliminated. Streamlines that passed through the MIMoSA were then saved binary .nii files, where 1 indicated that disease was present in that voxel, and 0 indicated either 1) that fascicle did not cross through that voxel or there was no disease in it.

    I was then able to calculate the "volume" of the disease in a fascicle (i.e. volume of the streamlines that were affected) by summing the # of 1s in the map. At the end, each individual had 77 single values that represented the volume of affected streamlines within each fascicle.

    Full fascicle volumes were also calculated and saved as .niis.

####  *Disease burden summary measures (done in this R script):*

    Having computed disease measures at the individual fascicle, I wanted to assess network effects for all future analyses. To do this, I calculated three summary measures per individual.

    1) Total disease : Sum of all 77 volume measures of disease, divided by the total volume of "healthy" fascicles (i.e. taking a sum of all of the full fascicle volumes). This yields a proportion of the overall burden of disease in the brain

    2) Depression network: Sum of all 19 volume measures of disease in the fascicles within the depression network, divided by the sum of total full fascicle volumes in the depression network.

    3) Nondepression network: Sum of all 58 volume measures of disease in the fascicles outside the depression network, divided by the sum of total full fascicle volumes in the nondepression network.

####  *Network effects:*

   Given that depression is so common in multiple sclerosis, we first wanted to check whether disease was diffuse across the brain, or whether there was network specificity (i.e. is MS random, or do MS lesions happen to strike fascicles that support the depression network, irrespective of diagnosis). To assess this, I took the following steps.

    1) Tests for normality (Shapiro–Wilk test) of proportion of disease in depression network and nondepression network. Both not normally distributed

    2) Used wilcoxon tests to compare disease (paired = T) between depression and nondepression network. As each individual contributed one measure for each network (i.e. 2 measures per person), paired was used.

####  *Diagnostic effects:*

  Per our inclusion/exclusion criteria, we were able to identify a subset of patients that had depression, as well as a subset who did not meet criteria per EMR review. We next wanted to assess whether patients with depression had more overall disease than patients without depression.

    First, we assessed whether lesion burden (i.e. the volume occupied by white matter disease), irrespective of which fascicles were impacted, differed between patients and controls.

    1) Tests for normality (Shapiro–Wilk test), volume of lesions not normally distributed.

    2) Used wilcoxon tests to compare lesion burden between depressed and non-depressed patients. As each person only contributed one measure, paired = F.

  Next, we assessed whether fascicle burden (i.e. total volume of disease divided by total volume of all fascicles) differed between diagnostic groups.

    1) Tests for normality (Shapiro–Wilk test), fascicle burden not normally distributed.

    2) Used wilcoxon tests to compare fascicle burden between depressed and non-depressed patients. As each person only contributed one measure, paired = F.

  We next aimed to see whether there was an interaction of Diagnosis and Network. Specifically, we hypothesized that patients with depression would have more disease specifically in the depression network.

     1) Model: diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)

  Given the somewhat arbitrary definition of depression network (25%/75%), we next assessed whether the relationship between diagnosis and network was continuous.

    1) For each fascicle, two values were computed

      a) Effect size (r) from the wilcoxon statistic comparing volume of disease in the fascicle between depressed vs nondepressed

      b) The volume of the overlap of that fascicle with the depression network

    2) A linear model relating the overlap of volume of the fascicle w/the depression network to the effect size from the depressed v nondepressed wilcoxon analysis.

    3) This relationship was permuted 10,000 w/boot (w/replacement)

    4) Plots were generated

####  *Sensitivity Analyses*

  Though the effect size by overlap analysis was reassuring, we wanted to further guarantee that our findings were not driven by arbitrary distinctions between the depression and non-depression networks. To assess this, we repeated the network analysis using tertiles and quintiles. LMER for assessing interactions was used.

    1) Tertiles:

      a) Depression network: Top Tertile
      b) Nondepression network: Bottom 2 Tertiles

    2) Quintiles:

      a) Depression network: Top Quintile
      b) Nondepression network: Bottom 4 Quintiles

## Results:

#### Demographics

232 with depression (depGroupVar == 2), 148 healthy (depGroupVar == 1)

                         Healthy          Depressed             p      test
  n                                148            232                     
  Race (%)         Caucasian       108 ( 73.0)    173 ( 74.6)   0.821     
                   Non-caucasian    40 ( 27.0)     59 ( 25.4)             
  Sex (%)          Female          117 ( 79.1)    199 ( 85.8)   0.117     
                   Male             31 ( 20.9)     33 ( 14.2)             
  Age (mean (SD))                46.57 (12.66)  48.75 (11.75)   0.089     
  Depression (%)   1               148 (100.0)      0 (  0.0)  <0.001     
                   2                 0 (  0.0)    232 (100.0)             
  PHQ2 (mean (SD))                0.00 (0.00)    0.58 (1.53)   <0.001     
  PHQ9 (mean (SD))                0.00 (NA)     10.43 (9.02)       NA

#### PROMIS Scores
                                               level   1              2              p      test
  n                                                       36             49                     
  Quality_of_Life (mean (SD))                           3.59 (1.02)    2.95 (0.99)    0.010     
  Physical_Health (mean (SD))                           2.86 (0.83)    2.41 (0.76)    0.019     
  Mental_Health_and_Mood (mean (SD))                    3.97 (0.82)    3.39 (0.99)    0.011     
  Social_Activities_Satisfaction (mean (SD))            3.83 (1.00)    2.84 (1.36)    0.001     
  Carrying_Out_Social_Activities (mean (SD))            3.71 (1.12)    2.93 (1.21)    0.007     
  Carrying_Out_Physical_Activities (mean (SD))          3.68 (1.06)    3.14 (1.00)    0.032     
  Emotional_Problems (mean (SD))                        3.67 (1.00)    3.00 (1.08)    0.011     
  **Fatique_Average (mean (SD))                           3.48 (1.25)    3.09 (1.14)    0.181  *   
  PostOp.PROMIS.Physical.Score (mean (SD))             13.15 (3.03)   11.67 (2.55)    0.032     
  PostOp.PROMIS.Mental.Score (mean (SD))               15.00 (2.92)   12.31 (3.31)    0.001

  This is actually great news - if depressed/nondepressed individuals differed on fatigue, you could make the case that of COURSE they'd feel worse as a consequence of their fatigue. However, that they weren't different here lends some specificity to the measures.

#### Depression Network Construction

dep_network_names_top_quartile_no_cranial_nerves
 [1] CC      AF_L    SLF2_R  SLF3_R  FAT_L   TR_S_R  SLF2_L  AF_R    TR_S_L  FAT_R   CS_S_R  CPT_F_R CPT_P_L ILF_R   CPT_P_R CS_S_L  PAT_R   MdLF_L  CBT_R

dep_network_names_bottom_3_quartiles_no_cranial_nerves
  [1] C_FPH_L C_FPH_R C_FP_L  C_FP_R  C_PH_L  C_PHP_L C_PHP_R C_PH_R  C_R_L   C_R_R   EMC_L   EMC_R   IFOF_L  IFOF_R  ILF_L   MdLF_R  PAT_L   SLF1_L  SLF1_R 
[20] SLF3_L  UF_L    UF_R    VOF_L   VOF_R   CB_L    CB_R    ICP_L   ICP_R   MCP     SCP     V       AR_L    AR_R    CBT_L   CPT_F_L CPT_O_L CPT_O_R CS_A_L 
[39] CS_A_R  CS_P_L  CS_P_R  CST_L   CST_R   DRTT_L  DRTT_R  F_L     F_R     ML_L    ML_R    OR_L    OR_R    RST_L   RST_R   TR_A_L  TR_A_R  TR_P_L  TR_P_R 
[58] AC

#### Network Effects

    - Disease within versus outside depression network: p = 2.2*10^-16, dep net prop = 0.400, non dep = 0.339, eff size 0.78 (large)

#### Diagnostic Effect

    - Volume of MIMoSA lesion burden: W = 14982, p=0.036; depressed = 15,871; nondepressed = 12,886, eff size 0.107 (small)

    - Proportion Volume of whole brain fascicle disease burden: W=14996, p=0.04, dep = 0.39, nondep = 0.34, eff size 0.11 (small)

#### Network x Diagnosis Interaction

    - Main effect of Diagnosis: T = 2.4, p = 0.015 (Depressed patient disease > Nondepressed patient disease)

    - Main effect of Network: T = 19.2, p <2 * 10^-16 (Depression Network more disease than nondepression network)

    - Interaction= T = 2.3, p = 0.02; finding driven by worse disease in depressed patients, within depression network

#### Effect Size x Fascicle overlap with depression network scatter

    - T = 2.4, p=0.02, adjusted R^2 = 0.06 (no boot)

#### Sensitivity Analyses

      1) Tertiles

        - Main effect of network (Depression > Nondepresison network): T=22.7, p = 2*10^-16

        - Main effect of diagnosis (Depressed > Nondepressed people):  T = 2.4, p = 0.015

        - Interaction of network and diagnosis:  T = 2.2, p = 0.029

      2) Quintiles

        - Main effect of network (Depression > Nondepresison network): T=23.0, p = 2*10^-16;

        - Main effect of diagnosis (Depressed > Nondepressed people): T = 2.4, p = 0.048

        - Interaction of network and diagnosis, T = 2.4, p = 0.018

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(boot)
library(boot.pval)
library(lsr)
library(sjPlot)
library(coin)
library(rstatix)
library(effectsize)
library(colormod)
library(ggpubr)
library(ggsignif)

homedir <- "/Users/eballer/BBL/msdepression/"

#functions
source(paste0(homedir,"/scripts/msdepression_functions.R"))

```

### Read in Data and do some preprocessing

```{r read_in_df_and_subset}

data <- read.csv(paste0(homedir, "/data/drugs_data/parsable_msdepression.csv"), sep = ",", header = TRUE) #n=16830, only has people with MS ICD10 code G35 and seen by MS provider

columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE")
                             #,"PHQ.2","PHQ.9")

#columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE","hemoglobin", "WBC","RBC","B12","FOLATE", "TSH", "RPR","CSFAPPEAR1","CSFAPPEAR4","CSFCOLOR1",       "CSFCOLOR4","CSFTUBE","CSFTUBE4","VitD","PHQ.2","PHQ.9")

##########################
### some preprocessing ###
##########################
#we need to keep scans from people seen by an MS provider (n = 17067) who have an MS diagnostic code (n=16,830)
#we make a bunch of columns from character to integer, binarize sex and race, and put date into a nice format YYYYMMDD, the %m%d%y indicates what it started out as

 data_empi_acc_f_phq <- data %>% 
   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = "")) %>%
   mutate(BEGIN_EXAM_DTTM = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% #make sure to actually save it as date object for later processing
   mutate(EMPI = as.factor(EMPI)) %>%
   mutate(On.Psych.Meds = ifelse(On.Psych.Meds == "True", 1, 0)) %>%
   mutate(On.Antidepressants = ifelse(On.Antidepressants == "True", 1, 0)) %>%
   mutate(Has.PHQ2 = ifelse(Has.PHQ2 == "True", 1, 0)) %>%
   mutate(Has.PHQ9 = ifelse(Has.PHQ9 == "True", 1, 0)) %>%
   mutate(Has.depdx = ifelse(grepl("F3", ICD10), 1, 0)) %>%
   mutate(PHQ.2_modsev_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 >= 3), 1, 0)) %>%
   mutate(PHQ.9_modsev_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 >= 10), 1, 0)) %>%
   mutate(PHQ.2_mild_dep_sxs = ifelse((!is.na(PHQ.2) & PHQ.2 > 0 & PHQ.2 < 3), 1, 0)) %>%
   mutate(PHQ.9_mild_dep_sxs = ifelse((!is.na(PHQ.9) & PHQ.9 > 0 & PHQ.9 < 10), 1, 0)) %>%
   mutate(PHQ.2_zero = ifelse((!is.na(PHQ.2) & PHQ.2 == 0), 1, 0)) %>%
   mutate(PHQ.9_zero = ifelse((!is.na(PHQ.9) & PHQ.9 == 0), 1, 0)) %>%
   mutate(dep_by_dx_phq = ifelse((Has.depdx | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(dep_by_dx_phq_antidep = ifelse((Has.depdx | On.Antidepressants | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(true_healthy = ifelse((!Has.depdx & !On.Psych.Meds & (PHQ.2_zero | PHQ.9_zero)), 1, 0)) %>%
   mutate(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds = ifelse(dep_by_dx_phq_antidep | true_healthy, 1, 0)) %>%
   rowwise(ACCESSION_NUM) %>%
   mutate(depGroupVar =
            sum(c(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds,dep_by_dx_phq_antidep))) %>%#you get an extra point if you are in the depression group, so the end result is dep = 2, healthy = 1, 0 for exclude
   ungroup() #n=3,737 unique people, n= 16,830 total

##########################################################
#             Read in Fascicle info             #
##########################################################

fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336, n unique = 958

#change empi to factor

fascicle_proportions$EMPI = as.factor(fascicle_proportions$EMPI)

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77, drop cranial nerves

#add suffix to each measure
fascicle_names_w_prop_suffix <- gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE, x = fascicle_names_no_cranial_nerves)
fascicle_names_w_vol_suffix <- gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE, x = fascicle_names_no_cranial_nerves)

########
#fascicle volumes in depression network
########

fascicle_volumes_dep <- read.csv(paste0(homedir, "/results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")

##########################################################
#             Read in Lesion Volumes                     #
##########################################################

# contains the volume of lesioned voxels, i.e. a sum of all the areas of each individual's binary mask that are 1

lesion_volumes <- read.csv(paste0(homedir, "/results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes"), sep = ",", header = TRUE) #n=958 unique, identical to fascicle unique

lesion_volumes$EMPI <- as.factor(lesion_volumes$EMPI)

##########################################################
#             Merge with data_empi              #
##########################################################
#merge whole data set with fascicle proportions
df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962, unique = 783

#merge with the df that has lesion volumes, 1 per perosn
df_demo_and_fascicles <- merge(df_demo_and_fascicles, lesion_volumes, by = c("EMPI", "EXAM_DATE")) #n = 2962

##########################################################
#  Add columns for volume and change proportion names    #
##########################################################

#read in fiber volume values
fiber_volume_values <- read.csv(paste0(homedir, "templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv"), sep=",", header = T)

#remove cranial nerves
fiber_volume_values_no_cranial_nerves <- fiber_volume_values %>% filter(fascicle %in% fascicle_names_no_cranial_nerves) %>%
  slice(order(factor(fascicle, levels = fascicle_names_no_cranial_nerves)))

#extract data frame that only includes the fascicle proportion lost
df_fascicles_prop <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)] 
df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)]

#multiply each row by the vector of total volumes to extract volume of disease within fascicle
df_fascicles_volumes <- sweep(x = df_fascicles_prop, MARGIN = 2,                        # Sweep with Complex STATS
                  STATS = fiber_volume_values_no_cranial_nerves$volume_full, FUN = "*") %>%
  mutate_if(is.numeric, round) #round all values

# change name of df_fascicle_volumes to have the suffix, and add a suffix for the proportion measures
names(df_fascicles_prop) <- fascicle_names_w_prop_suffix
names(df_fascicles_volumes) <- fascicle_names_w_vol_suffix

#combine back with data_fram
#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
df_demo_and_fascicles_prop_and_vol <- cbind(df_everything_else, df_fascicles_prop, df_fascicles_volumes)

```

```{r get_mimosa_QC_info}

#This section checks how many unique individuals had QA data, and then how many met MIMoSA QC. Some additional people lost at registration.

#1 Total unique individuals who also have MS - 890
#2 Total unique who had good MIMoSA - 786
#3 Total unique with bad MIMoSA - 890-786 = 104
#4 Total lost in streamline filtering - 3
#5 Final N going into inclusion/exclusion - 783

#read in mimosa data
mimosa_qc <- read.csv(paste0(homedir, "/data/melissa_martin_files/csv/mimosa_dataframe"), sep = ",", header = FALSE) #n=2840

#preprocess data frame to extract dates, empis, and score
empis_with_qc <- mimosa_qc %>% 
  mutate(EMPI = as.factor(gsub(x = V5, pattern = ".*sub-(.*)/ses-.*/.*", replacement="\\1", perl = TRUE))) %>%
  mutate(EXAM_DATE = gsub(x=V5, pattern = ".*/ses-(.{8})/.*", replacement = "\\1\\2\\3")) %>% 
  mutate(score = V3)

#merge with whole data frame to make sure we exclude people w/out MS
mimosa_qc_with_data_empi_acc_f_phq <- merge(data_empi_acc_f_phq, empis_with_qc, by = c("EMPI", "EXAM_DATE")) #n = 3592

#get list of unique patients, n = 890
mimosa_qc_with_data_empi_acc_f_phq_unique <- mimosa_qc_with_data_empi_acc_f_phq %>%
  group_by(EMPI) %>%
  arrange(EXAM_DATE) %>%
  slice(1) %>%
  ungroup #n=890

#get list of unique patients with good QC, n = 786
mimosa_unique_score_gre_75 <- mimosa_qc_with_data_empi_acc_f_phq %>%
  filter(score >= 75) %>%
  group_by(EMPI) %>%
  arrange(EXAM_DATE) %>%
  slice(1) %>%
  ungroup #n=786

#merge with fascicle proportions, we lose 3
mimosa_qc_with_data_empi_acc_f_phq_fascicle_props <- merge(mimosa_qc_with_data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n=2962

mimosa_qc_with_data_empi_acc_f_phq_fascicle_props_unique <- mimosa_qc_with_data_empi_acc_f_phq_fascicle_props %>% 
  group_by(EMPI) %>%
  arrange(EXAM_DATE) %>%
  slice(1) %>%
  ungroup #n=783
```

### PROMIS DATA HANDLING to link a person to their most proximal PROMIS

```{r promis_data_handling}
##########################################################
#            Read in PROMIS Data              #
##########################################################
promis_data <- readRDS(paste0(homedir, "/data/dac/promis_data_20220824_fixed_physical_health_cleaned_n1813_unique_indiv_n532_2015-2022.rds"))
promis_score_names <- names(promis_data[7:16])

df_people_with_promis_data <- df_demo_and_fascicles_prop_and_vol %>%
  filter(EMPI %in% promis_data$EMPI) #n=398

unique_empis_in_promis_group <- unique(df_people_with_promis_data$EMPI) #n=121

#go through each person w/promis data and append the columns to their row that reflects the closest collected PROMIS

#function, take an empi and date, returns the row of data w/promis scores
get_closest_promis_to_ms_exam_date <- function(subj_empi, subj_exam_date) {
  df_closest_promis <- promis_data %>%
    filter(EMPI == subj_empi) %>%
    mutate(distance_from_scan = abs(RECORDED_TIME-subj_exam_date)) %>%
    filter(abs(RECORDED_TIME-subj_exam_date)==min(abs(RECORDED_TIME-subj_exam_date))) %>%
    slice(1)
  return(df_closest_promis)
}

#make dataframe to store promis info
promis_info_closest_proximity <- data.frame(matrix(NA, nrow=dim(df_people_with_promis_data)[1], ncol=(dim(promis_data)[2] + 1)))
names(promis_info_closest_proximity) <- c(names(promis_data), "distance_from_scan")

for (line in 1:dim(df_people_with_promis_data)[1]) {
  #pull out subject empi and exam date
  subj_empi <- df_people_with_promis_data$EMPI[line]
  subj_exam_date <- df_people_with_promis_data$BEGIN_EXAM_DTTM[line]

  #returns whole line, promis scores, id, empi, everything
  closest_promis_data <- get_closest_promis_to_ms_exam_date(subj_empi = subj_empi, subj_exam_date = subj_exam_date)

  #store as new line
  promis_info_closest_proximity[line,] = closest_promis_data
}

#drop unnecessary columns c("DE_PAT_ID", "EMPI", "HUP_MRN", "SEX", "RACE")
promis_info_closest_proximity_just_scores <- promis_info_closest_proximity[,6:dim(promis_info_closest_proximity)[2]]

#combine with df_people_with_promis_data
df_people_with_promis_data_scores_added = cbind(df_people_with_promis_data, promis_info_closest_proximity_just_scores)

#summarize the distance from scans
summary(df_people_with_promis_data_scores_added$distance_from_scan/364) #mean 2.40 ys, median 2.10 years, range 0.003 y to 6.83 years

hist(df_people_with_promis_data_scores_added$distance_from_scan/364, 
     main = "Hisogram of years between Promis and Scan",
     xlab = "Years from Scan") #convert to years

```

### Demographics
```{r demographics}

#this will display a demographics table comparing age, race, sex between healthy and depressed groups, using the first presentation to treatment

df_good_mimosa_unique_dep_and_healthy <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique_dep_and_healthy)

#just good mimosa w/promis
df_people_with_promis_data_scores_added_unique <- df_people_with_promis_data_scores_added %>%
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms_w_promis(data_frame = df_people_with_promis_data_scores_added_unique)

```

### Define Depression Network by Volume of Overlap
```{r visualize_depression_network_volume}
###################################################
### Defining and Visualizing Depression Network ###
###################################################

#This chunk reads in a file where columns are fascicle, num_voxels_total, non_zero_voxels_in_dep_map, prop_in_mask
# by the end, the depression mask will be defined
# We consider fascicles in the depression network to be the ones that share the most VOLUME with the depression mask
# Of course, it is possible that huge fascicles contribute a ton to the depression network, but also the non-depression network
# And to get at specificity (i.e. which fibers most exclusively send information within the depression network), we could look at proportion of the fiber (volume that intersects / total volume of the fascicle). But then you have a situation where tiny fibers get lots of credit. We have actually tried this as well and the findings aren't that different. 
#But for the sake of this paper, will do volume because the question is all volume based (volume of lesions to volume of fascicles)

######## Histograms for depression network
#first set the values to be the volumes
fascicle_volumes_dep_network_no_cranial_nerves <- fascicle_volumes_dep %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

#set NAs to zero
fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map[is.na(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map)] = 0

#sort based on overlap with depression depression map
fascicle_volumes_dep_network_no_cranial_nerves$fascicle <- factor(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, levels = fascicle_volumes_dep_network_no_cranial_nerves$fascicle[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map)])

#figure out how many fascicles are in top quartile
quartile <- round((dim(fascicle_volumes_dep_network_no_cranial_nerves)[1] / 4),0)

#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take names of the top quartile
dep_network_names_top_quartile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile]

write.table(dep_network_names_top_quartile_no_cranial_nerves, paste0(homedir, "/results/dep_network_names_top_quartile_no_cranial_nerves_by_vol.txt"), row.names = F, col.names = F, quote = F)

df_for_making_binary_color_maps <- fascicle_volumes_dep_network_no_cranial_nerves %>% dplyr::select(fascicle, non_zero_voxels_in_dep_map)

write.table(df_for_making_binary_color_maps, paste0(homedir, "/results/volume_of_overlap_w_dep_net_n77.csv"), row.names = F, col.names = F, quote = F, sep = ",")

#get names of bottom 3 quartiles

#get a network that includes all of the fascicles NOT in dep 25% network
dep_network_names_bottom_3_quartiles_no_cranial_nerves <- fascicle_volumes_dep_network_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves)]

#### Get sums of the volumes of the fascicles in each network, and total sum

#top quartile, depression network
dep_network_total_volume_top_quartile_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves)])

#bottom 3 quartiles, nondepression network
bottom_75_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves)])

#### Calculate total volume in all non-cranial_nerve_network
total_volume_all_full_fascicles <- dep_network_total_volume_top_quartile_no_cranial_nerves + bottom_75_total_volume_no_cranial_nerves

###

plot_all_fascicle_volumes_dep_network_no_cranial_nerves <-ggplot(data = fascicle_volumes_dep_network_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map))+
  ggtitle("Volume Overlap per Fascicle 3_09") + 
  geom_bar(stat="identity", fill="006300") +
  xlab("Fascicle") +
  ylab("Overlap with Depression Mask") +
  ylim(0,40000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicle_volumes_dep_network_no_cranial_nerves)

#save pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/plot_all_fascicle_volumes_dep_network_no_cranial_nerves.pdf"))
  plot_all_fascicle_volumes_dep_network_no_cranial_nerves
dev.off()

### Change names to add suffix
#replace each fascicle with vol suffix
dep_network_names_top_quartile_no_cranial_nerves <- dep_network_names_top_quartile_no_cranial_nerves %>% 
  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix

dep_network_names_bottom_3_quartiles_no_cranial_nerves <- dep_network_names_bottom_3_quartiles_no_cranial_nerves %>%
   gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix

```

### Compare overall volume of the lesions, irrespective of fascicle involvement, between diagnostic groups

```{r compare_overall_volume_of_lesions_dep_vs_non_dep}

#separate out only people in depression group, and take first instance
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# Assessing Normality w/histograms and Shapiro_Wilk test for normality
hist(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])
non_dep_lesion_vol_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]) #W=0.77, p = 8.2 * 10^-14, NON-NORMAL

hist(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])
dep_lesion_vol_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]) #W=0.80, p = 2.2 * 10^-16, NON-NORMAL

dep_vs_healthy_wilcoxon_lesion_volume_results <-  wilcox.test(dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$volume_of_mimosa_lesions[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired = F) #W = 14982, p=0.036; mean depressed = 15,871; mean nondepressed = 12,886

dep_vs_healthy_wilcoxon_lesion_volume_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = volume_of_mimosa_lesions ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000) #eff size 0.107 (small)

#extract overall volume and diagnosis
df_lesion_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
  dplyr::select(volume_of_mimosa_lesions,depGroupVar) %>%
  filter(depGroupVar != 0) %>% 
  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))

violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, fill=depGroupVar)) +#, fill=depGroupVar)) + 
  geom_violin() + 
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
  ylab("Volume of Lesions") +
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
print(violin_lesion_volume) #p=0.067

#save pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/violin_lesion_volume.pdf"))
  violin_lesion_volume
dev.off()

```

### Compare between networks, diagnostic groups, and interaction

```{r depression_vs_healthy_overall_volume_of_fascicles_affected}
## compare disease within and outside depression network, between diagnoses, and interactions

###############################
### Constructing Data Frame ###
###############################

#make a data frame for only unique individuals, and get means of overall volume of disease across all fascicles, and within/outside dep net
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>%

  #sum of total volume lost and proportion of volume lost 
  mutate(sum_fascicle_vol_lost = rowSums(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  mutate(proportion_volume_lost_per_total_network_size_sum = sum_fascicle_vol_lost/total_volume_all_full_fascicles) %>%

  #sum of total volume lost in depression network and proportion of volume lost in depression network (volume lost/network volume size) 
  mutate(sum_fascicle_vol_lost_indepnet_top_quartile = 
           rowSums(across(all_of(dep_network_names_top_quartile_no_cranial_nerves)))) %>%
  mutate(proportion_volume_lost_dep_net_by_network_size_sum = sum_fascicle_vol_lost_indepnet_top_quartile /dep_network_total_volume_top_quartile_no_cranial_nerves) %>%

  #sum of total volume lost in nondepression network and proportion of volume lost in depression network (volume lost/network volume size) 
  mutate(sum_fascicle_vol_lost_nondep_bottom_3_quartiles = 
           rowSums(across(all_of(dep_network_names_bottom_3_quartiles_no_cranial_nerves)))) %>%
  mutate(proportion_volume_lost_nondep_net_by_network_size_sum = sum_fascicle_vol_lost_nondep_bottom_3_quartiles /bottom_75_total_volume_no_cranial_nerves) %>%

  ungroup()

###################

##### Assessing Normality w/histograms and Shapiro_Wilk test for normality ####

### Test for normality within each network, within each diagnosis

disease_overall_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum) #W=0.97, p = 4.1 * 10^-7, NON-NORMAL

disease_dep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum) #W=0.97, p = 2.0 * 10^-7, NON-NORMAL

disease_nondep_net_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum) #W=0.97, p = 2.6 * 10^-7, NON-NORMAL

disease_dep_dx_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]) #W=0.97, p = 0.0003 NON-NORMAL

disease_nondep_dx_shapiro <- shapiro.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]) #W=0.96, p = 0.0001 NON-NORMAL

## All are non-normal, will use wilcoxon for comparisons

### Within/outside depression network, V = 65897, p = 2.2*10^-16, mean dep net prop = 0.41, mean non dep = 0.31
msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum,paired = TRUE)

#within nondepressed group, V=9661, p <2.2*10^-16
msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1],paired = TRUE)

#within depressed group, V= 25143, p<2.2*10^-16
msdisease_within_versus_outside_depression_network_dep_group_sums_scaled_by_size_wilcoxon <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2],paired = TRUE)

### Diagnostic effects

#proportion volume lost over whole network = W=14996, p=0.04, mean dep = 0.39, mean nondep = 0.34
proportion_volume_lost_over_whole_brain_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_per_total_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)

#effsize 0.11 (small)
proportion_volume_lost_over_whole_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_per_total_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)

#proportion volume lost within depression network, W = 14969, p-value = 0.04; dep = 0.44, nondep = 0.38
proportion_volume_lost_dep_net_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)

#effsize 0.11 (small)
proportion_volume_lost_over_dep_net_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_dep_net_by_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)

#proportion volume lost outside depression network W = 15134, p-value 0.051; mean dep = 0.32, mean nondep = 0.29
proportion_volume_lost_nondep_net_dep_vs_healthy <- wilcox.test (dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], paired=FALSE)

#effsize 0.10 (0.099) (small)
proportion_volume_lost_over_nondep_net_brain_dep_vs_healthy_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = proportion_volume_lost_nondep_net_by_network_size_sum ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)

### Interaction Network x Diagnosis

df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

#effect size within vs outside dep network, eff size = 0.78 (large)
msdisease_within_versus_outside_depression_network_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions, formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)

#effect size dep vs nondep network within nondepressed, eff size = 0.75 (large)
msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "Nondepressed"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)

#effect size dep vs dep network within depressed, eff size = 0.80, large
msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "Depressed"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)

#diagnosis: Dep > nondep T = 2.4, p = 0.015; Network dep>nondep = T=19.2, p = 2*10^-16; intx = T = 2.3, p = 0.020
diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)
summary(diagnosis_by_network_interaction_lmer)

```
### Plot interaction
``` {r plots_for_dx_net_intx}
########## PLOTS #############

#### color scheme
Diagnosis_colors <- c("#AB19E6", "#1954E6")
Diagnosis_colors_faded <- as.vector(sapply(Diagnosis_colors, function(color) {decrease_saturation(color, 0.3)})) 
Network_colors <- c("#ff00ff", "#3df2f2")
Network_colors_faded <- as.vector(sapply(Network_colors, function(color) {decrease_saturation(color, 0.3)}))

#label scheme
Network_labels <- c("Depression Network", "Nondepression Network")
Network_labels_wrapped <- c("Depression\nNetwork", "Nondepression\n Network")
Diagnosis_labels <- c("Depressed", "Nondepressed")

#Within vs outside network
dep_net_group <- c(rep("Depression Network", times = dim(dep_and_healthy_groups_for_ICD_analysis)[1]), rep("Nondepression Network",times =dim(dep_and_healthy_groups_for_ICD_analysis)[1]))
net_values <- c(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum)

newdf <- data.frame(dep_net_group, net_values)

newdf <- newdf %>%
  mutate( dep_net_group=factor(dep_net_group,levels=c( "Nondepression Network", "Depression Network")))

violin_within_vs_outside_network_sum_proportion_scaled_by_network_size <- ggplot(newdf, aes(x=dep_net_group, y=net_values, fill=dep_net_group)) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  ggtitle(paste0("Volume of Fascicle Disease Burden")) +
  ylab("Proportion of network affected") +
  ylim(0,1) + 
  scale_fill_manual(values=c("#80ffff", "#ff40ff")) +
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 16, colour = "black"), axis.title.y  = element_blank(), 
        axis.title.x = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=4/7,
        panel.background = element_rect(colour = "white", fill = "white")) + 
  coord_flip()

print(violin_within_vs_outside_network_sum_proportion_scaled_by_network_size) #

#save pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/violin_within_vs_outside_network_sum_proportion_scaled_by_network_size.pdf"))
  violin_within_vs_outside_network_sum_proportion_scaled_by_network_size
dev.off()

#diagnosis

df_proportion_network_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
dplyr::select(proportion_volume_lost_per_total_network_size_sum, depGroupVar) %>%
  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))

violin_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis, aes(x=depGroupVar, y=proportion_volume_lost_per_total_network_size_sum, fill=depGroupVar)) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "NonDepressed", "2" = "Depressed")) + 
  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
  ylab("Proportion (sum total volume lost/Net Vol) ") +
  ylim(0,1) + 
  scale_fill_manual(values=Diagnosis_colors) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(violin_dep_vs_nondep_scaled_by_network_size) #

#save pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/violin_dep_vs_nondep_scaled_by_network_size.pdf"))
  violin_dep_vs_nondep_scaled_by_network_size
dev.off()

# bar graph instead of violin

df_proportion_network_volume_and_diagnosis_summary <- df_proportion_network_volume_and_diagnosis %>%
  group_by(depGroupVar) %>%
  summarise(
    mean = mean(proportion_volume_lost_per_total_network_size_sum),
    sd = sd(proportion_volume_lost_per_total_network_size_sum),
    n = n(),
    se = sd / sqrt(n)
  )

bar_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis_summary, aes(x=depGroupVar, y=mean, fill=depGroupVar)) +
  geom_bar(stat = "identity", fill=Diagnosis_colors_faded, color = Diagnosis_colors, size = 1.5, width = 0.8) +  
  geom_errorbar(aes(ymin=mean-se,ymax=mean+se), width=0.2, size=1.5, color = 'black') + #, position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
  scale_x_discrete(labels=c("Non-depressed" = "Nondepressed", "Depressed" = "Depressed")) + 
#  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
  ylab("Proportion of Whole Brain White Matter Volume Affected") +
  ylim(0,0.6) + 
  theme(plot.title = element_text(size=14, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=14),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(bar_dep_vs_nondep_scaled_by_network_size) #

#save pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/bar_dep_vs_nondep_scaled_by_network_size.pdf"))
  bar_dep_vs_nondep_scaled_by_network_size
dev.off()

#plot interaction with ggplot rather than plot_model

df_interaction <- expand.grid(Network = levels(df_empi_dx_network_proportions$Network), Diagnosis = levels(df_empi_dx_network_proportions$Diagnosis), subject = "8450972724") # just pick a random person

model_data <- get_model_data(model = diagnosis_by_network_interaction_lmer, type = "int")

df_interaction_fit_upper_lower <- cbind(df_interaction, model_data$predicted, model_data$conf.high, model_data$conf.low)

### line

interaction_line <- ggplot(data = df_interaction_fit_upper_lower, aes(x = Network, y = model_data$predicted, color = Diagnosis)) +
  geom_line(aes(group = Diagnosis), size = 1.5) +
  scale_color_manual(values=Diagnosis_colors, labels=Diagnosis_labels)+ 
  geom_point(size = 3) + 
 # geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), width = 0.2,size = 1, color='black') +
  scale_x_discrete(labels=Network_labels) +
  ylim(0,0.6) + 
    xlab("Diagnosis") + 
    ylab("Proportion Affected") + 
    theme(axis.text.x = element_text(size=16, color="black"),
          axis.text.y = element_text(size=16, color="black"),
          axis.title.x = element_blank(), #element_text(size =16), 
          axis.title.y = element_text(size =16), 
          title = element_text(size=18),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          aspect.ratio = 7/5,
          panel.background = element_rect(colour = "white", fill = "white"),
          legend.key=element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 14), legend.position = "top")

#save as pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_line_graph_range0to06_point.pdf"))
  interaction_line
dev.off()

#geom_point
interaction_ggplot_point <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network, group = interaction(Diagnosis, Network))) + 
    geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
    geom_point(position = position_dodge(width=0.5), size = 5) + 
    scale_color_manual(values=Network_colors, labels = c("Depression Network", "Nondepression Network")) + 
    ylim(0,1) + 
    xlab("Diagnosis") + 
    ylab("Proportion Affected") + 
    theme(axis.text.x = element_text(size=16, color="black"),
          axis.text.y = element_text(size=16, color="black"),
          axis.title.x = element_blank(), #element_text(size =16), 
          axis.title.y = element_text(size =16), 
          title = element_text(size=18),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          aspect.ratio = 5/3,
          panel.background = element_rect(colour = "white", fill = "white"),
          legend.key=element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 14), #,
          legend.position = "top") #+ 
 #   ggtitle("Diagnosis x Depression Network Disease")
print(interaction_ggplot_point)

#save as pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0to1_point.pdf"))
  interaction_ggplot_point
dev.off()

#geom_point 0.2-0.5
interaction_ggplot_point_02_06 <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network, group = interaction(Diagnosis, Network))) + 
    geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
    geom_point(position = position_dodge(width=0.5), size = 5) + 
    scale_color_manual(values=Network_colors, labels = c("Depression Network", "Nondepression Network")) + 
    ylim(0.2,0.6) + 
    xlab("Diagnosis") + 
    ylab("Proportion Affected") + 
    theme(axis.text.x = element_text(size=16, color="black"),
          axis.text.y = element_text(size=16, color="black"),
          axis.title.x = element_blank(), #element_text(size =16), 
          axis.title.y = element_text(size =16), 
          title = element_text(size=18),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          aspect.ratio = 5/3,
          panel.background = element_rect(colour = "white", fill = "white"),
          legend.key=element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 14), #,
          legend.position = "top") #+ 
 #   ggtitle("Diagnosis x Depression Network Disease")
print(interaction_ggplot_point_02_06)

#save as pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0_2to0_6_point.pdf"))
  interaction_ggplot_point_02_06
dev.off()

#geom_bar
interaction_ggplot_bar <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network,  group = interaction(Diagnosis, Network))) + 
    geom_col(position = position_dodge(width=0.6), fill = c(Network_colors_faded, Network_colors_faded), color = c(Network_colors, Network_colors), width = 0.5, size = 1.5) + 
    geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.6), width = 0.2,size = 1.5, color='black') +
    scale_fill_manual(values=Network_colors) + #, labels = c("Depression Network", "Nondepression Network"))+ 
    ylim(0,0.6) + 
    xlab("Diagnosis") + 
    ylab("Proportion of White Matter Network Affected") + 
    theme(axis.text.x = element_text(size=14, color="black"),
          axis.text.y = element_text(size=14, color="black"),
          axis.title.x = element_blank(), #element_text(size =16), 
          axis.title.y = element_text(size =14), 
          plot.title = element_text(size=16, hjust = 0.5),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          aspect.ratio = 7/5,
          panel.background = element_rect(colour = "white", fill = "white"))+

    geom_signif(
      y_position = c(0.48, 0.48, 0.53, 0.58), 
      xmin = c(0.85, 1.85, 0.85, 1.15),
      xmax = c(1.15, 2.15, 1.85, 2.15),
      annotation = c("***", "***", "*", "NS"), 
      tip_length = 0.03, color = "black",
      size = 1.1, textsize = 5) #+
  #  ggtitle("Diagnosis x Depression Network Disease")
print(interaction_ggplot_bar)

#save as pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0to06_bar.pdf"))
  interaction_ggplot_bar
dev.off()

### wrap

diagnosis_and_interaction_bar_combined <- ggarrange(bar_dep_vs_nondep_scaled_by_network_size,interaction_ggplot_bar, 
                                                labels = c("A", "B"),
                                                ncol = 2, nrow = 1)

print(diagnosis_and_interaction_bar_combined)
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_and_intx_bargraph_range0to06.pdf"))
  diagnosis_and_interaction_bar_combined

diagnosis_and_interaction_all_combined <- ggarrange(bar_dep_vs_nondep_scaled_by_network_size, interaction_line,interaction_ggplot_bar, 
                                                labels = c("A", "B", "C"),
                                                ncol = 3, nrow = 1)

pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_intx_line_and_intx_bargraph_range0to06.pdf"))
  diagnosis_and_interaction_all_combined
dev.off()
```

### Effect size regression (effect size x overlap with dep net)
``` {r effect_size_regression}
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0)  %>% 
 # mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# 
#depGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_effsize <- sapply(fascicle_names_w_vol_suffix, function(fascicle) 
{

  text_for_eval <- paste0("rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = ", fascicle, " ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = \"perc\", nboot = 1000)$effsize")
  eval(parse(text=text_for_eval))
})
names(fascicle_dep_vs_healthy_effsize) <- fascicle_names_w_vol_suffix

#convert to data frame
df_effsize_values <- as.data.frame(fascicle_dep_vs_healthy_effsize)

write.table(x = df_effsize_values, file = paste0(homedir, "results/fascicle_and_wilcox_effect_size_for_volxdx_analysis_n77.csv"), row.names = T, col.names = F, quote = F, sep = ",")

```

### Plot effect size x depression net overlap
``` {r plot_scatter}
#######################
### plot differences ##
#######################

#combine data frames, column 1 has names of fascicles, column 2 has Volume of overlap with depression network, column 3 has t-test value of dep vs healthy unc, 4th dep vs healthy fdr corrected within that fascicle

df_fascicle_dep_mask_volume_overlap_effsize_value <- data.frame(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, df_effsize_values)
names(df_fascicle_dep_mask_volume_overlap_effsize_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_effsize")

# plot fascicle overlap x Eff size value
#statistic, p=0.02, T = 2.4, eff size (adjusted R^2 = 0.06)
dep_overlap_x_dep_vs_healthy_vol_effsize <- lm(dep_vs_healthy_effsize ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_effsize_value )
summary(dep_overlap_x_dep_vs_healthy_vol_effsize)

scatter_effsize_val_by_dep_net_overlap_log <- ggplot(data=df_fascicle_dep_mask_volume_overlap_effsize_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_effsize, color=dep_vs_healthy_effsize)) +
  geom_point(size=2) + 
  scale_color_gradient(low = "#FFFFFF", high = "#ff00ff") +
  geom_smooth(formula = y ~ x, method=lm, color="black")+
  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(0, 10, 100, 1000, 10000, 100000)) + 
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect Size (r)") + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text = element_text(size=16, color = "black"),
        axis.title = element_text(size=18, color = "black"),
        axis.line = element_line(color="black", size=1.3),
        panel.background = element_blank(),
        legend.position = "none",
        aspect.ratio = 4/4)
print(scatter_effsize_val_by_dep_net_overlap_log)

#save as pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/scatter_effsize_val_by_dep_net_overlap_log.pdf"))
  scatter_effsize_val_by_dep_net_overlap_log
dev.off()

df_for_making_colors <- df_fascicle_dep_mask_volume_overlap_effsize_value %>% dplyr::select(fascicle, dep_vs_healthy_effsize)
write.table(x = df_for_making_colors, file = paste0(homedir, "results/fascicle_and_effect_size_for_volxdx_analysis_n77.csv"), row.names = F, col.names = F, quote = F, sep = ",")

```

### Sensitivity Analyses

```{r sensitivity analyses}

## to check different splits of the networks

tertile <- round((dim(fascicle_volumes_dep_network_no_cranial_nerves)[1] / 3),0)
quintile <- round((dim(fascicle_volumes_dep_network_no_cranial_nerves)[1] / 5),0)

## get names of each network, and add suffix
dep_network_names_top_tertile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:tertile]  #n=26

dep_network_names_bottom_tertiles_no_cranial_nerves <- fascicle_volumes_dep_network_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves)]

dep_network_names_top_quintile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quintile]  #n=15

dep_network_names_bottom_quintiles_no_cranial_nerves <- fascicle_volumes_dep_network_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves)]

### get volumes of each network
top_tertile_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves)])

bottom_tertiles_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_tertiles_no_cranial_nerves)])

top_quintile_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves)])

bottom_quintiles_total_volume_no_cranial_nerves<-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_quintiles_no_cranial_nerves)])

## change suffix to _vol

dep_network_names_top_tertile_no_cranial_nerves <- dep_network_names_top_tertile_no_cranial_nerves %>% 
  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

dep_network_names_bottom_tertiles_no_cranial_nerves <- dep_network_names_bottom_tertiles_no_cranial_nerves %>% 
  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

dep_network_names_top_quintile_no_cranial_nerves <- dep_network_names_top_quintile_no_cranial_nerves %>%
  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

dep_network_names_bottom_quintiles_no_cranial_nerves <- 
dep_network_names_bottom_quintiles_no_cranial_nerves %>% 
  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

### make new columns
dep_and_healthy_groups_for_ICD_analysis <- dep_and_healthy_groups_for_ICD_analysis %>%

  #tertile and quintiles
  mutate(sum_fascicle_vol_lost_top_tertile = 
           rowSums(across(all_of(dep_network_names_top_tertile_no_cranial_nerves)))) %>%
  mutate(proportion_volume_lost_top_tertile_by_network_size_sum = sum_fascicle_vol_lost_top_tertile /top_tertile_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_bottom_tertiles = 
           rowSums(across(all_of(dep_network_names_bottom_tertiles_no_cranial_nerves)))) %>%
  mutate(proportion_volume_lost_bottom_tertiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_tertiles /bottom_tertiles_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_top_quintile = 
           rowSums(across(all_of(dep_network_names_top_quintile_no_cranial_nerves)))) %>%
  mutate(proportion_volume_lost_top_quintile_by_network_size_sum = sum_fascicle_vol_lost_top_quintile /top_quintile_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_bottom_quintiles = 
           rowSums(across(all_of(dep_network_names_bottom_quintiles_no_cranial_nerves)))) %>%
  mutate(proportion_volume_lost_bottom_quintiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_quintiles /bottom_quintiles_total_volume_no_cranial_nerves)

#### tertile/quintile

##### within versus outside network
#V=67727, p=2.2*10^16, top tertile prop vol affected = 0.42, bottom 2 tertiles = 0.28
msdisease_within_versus_outside_depression_network_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum,paired = TRUE)

#p=2.2*10^-16, top quintile prop vol affected = 0.43, bottom 4 quintiles = 0.30
msdisease_within_versus_outside_depression_network_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum,paired = TRUE)

#################
## Interaction ##
#################

#### tertile
df_empi_dx_network_proportions_tertiles <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_top_tertile_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_bottom_tertiles_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

#diagnosis: Dep > nondep T = 2.4, p = 0.015; Network dep>nondep = T=22.7, p = 2*10^-16; intx = T = 2.19, p = 0.029
diagnosis_by_network_interaction_lmer_tertile <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions_tertiles)
summary(diagnosis_by_network_interaction_lmer_tertile)

#### quintile
df_empi_dx_network_proportions_quintiles <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_top_quintile_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_bottom_quintiles_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

#diagnosis: Dep > nondep T = 2.4, p = 0.015; Network dep>nondep = T=23.0, p = 2*10^-16; intx = T = 2.37, p = 0.018
diagnosis_by_network_interaction_lmer_quintile <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions_quintiles)
summary(diagnosis_by_network_interaction_lmer_quintile)

```